---
title: "iNEXT_tutorial"
author: "Holden Jones"
date: "2024-09-28"
output: html_document
---

# PRACTICE RUNNING iNEXT
- using their tutorial
https://github.com/JohnsonHsieh/iNEXT
```{r}
library(iNEXT)
library(ggplot2)
```

# spider
Spider data (datatype = "abundance")
Description
The data include spider species abundances in two canopy manipulation treatments 
(Girdled and Logged) of hemlock trees (Ellison et al. 2010, Sackett et al. 2011). 
In the Girdled treatment site, there were 26 species among 168 individuals; 
in the Logged treatment site, there were 37 species among 252 individuals.

The format for each site is a list of species abundances:
A list of 2 vectors
$ Girdled: num [1:26] 46 22 17 15 15 9 8 6 6 4 ...
$ Logged : num [1:37] 88 22 16 15 13 10 8 8 7 7 ...
```{r}
data(spider)



out <- iNEXT(spider, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# Sample-size-based R/E curves, separating plots by "Assemblage"
ggiNEXT(out, type=1, facet.var="Assemblage")

# Sample-size-based R/E curves, separating plots by "Order.q"
ggiNEXT(out, type=1, facet.var="Order.q")

# To link the sample-sized and coverage-based sampling curves, 
#it would be informative to first examine the sample completeness curve using 
#the following command:
ggiNEXT(out, type=2)

# The following commands return the coverage-based R/E sampling curves. The argument facet.var=“Assemblage” in the ggiNEXT() function creates a separate plot for each site, as shown below:
ggiNEXT(out, type=3, facet.var="Assemblage")
```








