---
title: "ranas_basic_old"
author: "Holden Jones"
date: '2023-02-02'
output: html_document
---

Goal is to explore the ranas, individual data from my FCAT cacao project. The
questions I'm most interested in looking at are: how do abundance and diversity
vary across cacao types, in comparison to reference forest and abandoned 
plantations. I'm also interested in seeing how different environmental variables
either related to the sampling (weather, temp, humidity), or the habitat type 
(leaf coverage, vegetation coverage, canopy closure, proximity to forest)
influence both abundance and diversity. Use this script to explore the data,
once I find avenues for future analyses, break those into separate scripts.

# Cacao ranas explore

load packages, data
ranas - individual capture data
environmental - sampling environmental data

```{r load}
library(labdsv)
library(tidyverse)
library(vegan)

ranas <- read_csv("data/ranas_05_09_23.csv")
enviromental <- read_csv("data/environment_04_08_23.csv")
site_data <- read_csv("data/site_data.csv")

new <- ranas %>%
  filter(Recap == "N")
```


--------------------------------------------------------------------------------
# How many individuals have we caught?

simple ind caught for each site
*12/07/22 1114 total ind*
```{r ind_by_site}
# simple ind caught for each site - just look at number new ind caught
new %>%
  group_by(Sitio) %>%
  summarize(count = n_distinct(Numero)) %>%
  ggplot(aes(Sitio, count)) +
  geom_col()
```

how has number of ind changed over time?
*like this fig so far, but need to adjust axes, theme*
```{r ind_over_time}
new_test <- new %>%
  group_by(Fecha) %>%
  mutate(new_ind = as.numeric(max(Numero)))

# this is close, need to adjust axis ticks
new_test %>%
  ggplot(aes(x = factor(Fecha, level = c(
    '01/09/22', '02/09/22', '03/09/22', '05/09/22', '06/09/22',
    '07/09/22', '08/09/22', '09/09/22', '10/09/22', '12/09/22',
    '13/09/22', '14/09/22', '15/09/22', '17/09/22', '19/09/22',
    '03/10/22', '05/10/22', '07/10/22', '08/10/22', '10/10/22', 
    '12/10/22')), 
    y = new_ind)) +
  geom_point() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(x = 'Sampling Nights', y = 'Cumulative Unique Individuals')
```


--------------------------------------------------------------------------------
# How many species have we caught?

how many species have we found so far?
*12/07/22 30 species so far*
*note using SHJ ID, ultimately use FVE ID, but need him to go through them*
```{r}
new %>%
  summarize(count = n_distinct(`Final ID`)) ## WATCH OUT! 5/9/23
## new update(?) means that can no longer put these dashes? now have to use
## ` instead of ' or " WTF! that's crazy

## this will work though
new %>%
  count(`Final ID`)
```

how many new species by each sampling effort?

dad's take:
```{r cum_species}
ranas$sample_date <- as.Date(ranas$Fecha, format = "%d/%m/%y")

species_by_day <- ranas %>%
  group_by(sample_date) %>%
  summarize(species = list(sort(unique(`SHJ ID`))),
            no_species = n_distinct(`SHJ ID`))
 
species_by_day %>%
  ggplot(aes(sample_date, no_species)) +
  geom_col()

# accumulates values in a str column and creates two new columns in dataframe
# this assumes that dataframe is already sorted
# might also assume that dataframe is the result of a group_by? (ie the values are contained in a list)
# 1) cum_ + column_name - comma separated string of unique values
# 2) no_cum_ + column_name
accumulate_column <- function(df, column_name) {
  cum_items = list()   # create empty list to hold results
 
  # loop over each row in the dataframe
  for(i in 1:nrow(df)) {
    cum_items <- append(cum_items, df[i, column_name])     # add current row's items to results
    cum_items <- purrr::flatten(cum_items)      # flatten items to maintain a single list (not a list of lists)
    cum_items <- unique(unlist(cum_items))                 # keep results unique
   
    # add current accumulated results to row
    df[i, paste("cum_", column_name, sep = "")] <- paste(sort(unlist(cum_items)), 
                                                         collapse = ",")
    df[i, paste("no_cum_", column_name, sep = "")] <- length(cum_items)
  }
 
  return(df)
}

cum_species_by_day <- accumulate_column(species_by_day, "species")

cum_species_by_day %>%
  ggplot(aes(sample_date, no_cum_species)) +
  geom_point()

cum_species_by_day %>%
  count(`cum_species`)
```


--------------------------------------------------------------------------------
# How does abundance vary with habitat type?

simplest would be to make boxplot for Type ~ Individuals
here for C vs. N comparison
```{r}
# simple boxplot only for unique individuals
new %>%
  filter(Tipo == "N" | Tipo == "C") %>% # filtering just C, N b/c sampling effort
  group_by(Tipo) %>%
  summarize(count = n_distinct(Numero)) %>%
  ggplot(aes(Tipo, count)) +
  geom_col()
# wow! this shows basically no difference in number of new ind. C vs N

# what if we include recaptures? overall number of captures comparison
ranas %>%
  filter(Tipo == "N" | Tipo == "C") %>% # filtering just C, N b/c sampling effort
  group_by(Tipo) %>%
  summarize(count = n_distinct(Numero)) %>%
  ggplot(aes(Tipo, count)) +
  geom_col()
# interesting, seems to be a difference here, more overall captures in C vs. N
```

here for B vs. V comparison
```{r}
# with equal sampling effort, seems to be more ind in viejo than bosque
new %>%
  filter(Tipo == "B" | Tipo == "V") %>% 
  group_by(Tipo) %>%
  summarize(count = n_distinct(Numero)) %>%
  ggplot(aes(Tipo, count)) +
  geom_col()

# also overall more captures in viejo than bosque
ranas %>%
  filter(Tipo == "B" | Tipo == "V") %>%
  group_by(Tipo) %>%
  summarize(count = n_distinct(Numero)) %>%
  ggplot(aes(Tipo, count)) +
  geom_col()
```

## Thoughts - use mark recapture
Eventually will want to use mark recapture data to create more accurate 
abundance estimates. The issue is that we don't have individual marks, we have
one generic mark for site. So instead of having captures for each individual
over time, we will have a proportion of recaptured / total captures for each
sampling effort, with the number of recaptures increasing over time.

I think it will be helpful to work through the Scinax data with Miriam as a 
simpler introduction to this issue. Ultimately expect to do the same for each
species, perhaps only for most common - priach and maybe another few? Could just
use minimum number for other less common species.


--------------------------------------------------------------------------------
# How does richness vary with habitat type?

start with simple species richness, will be easiest thing to look at
boxplot of species richness for each comparison (C/N, B/V)
```{r}
# shows slightly higher diversity in Nacional, but doesn't seem like much
ranas %>%
  filter(Tipo == "N" | Tipo == "C") %>%
  group_by(Tipo) %>%
  summarize(count = n_distinct(`SHJ ID`)) %>%
  ggplot(aes(Tipo, count)) +
  geom_col()

# for b vs v comparison
ranas %>%
  filter(Tipo == "B" | Tipo == "V") %>%
  group_by(Tipo) %>%
  summarize(count = n_distinct(`SHJ ID`)) %>%
  ggplot(aes(Tipo, count)) +
  geom_col()
# substantially more species richness in bosque compared to viejo!
```

## Thoughts - use better diversity index
For exploratory analyses it's fine to start with species richness, but ultimately
will want to use a better one. Need to review the different options for calculating
species diversity.
- simpson's - prob 2 ind are same species
- shannon-weiner - certainty of identity of random ind
- evenness - how similar in number are species?


--------------------------------------------------------------------------------
# Survey

read data
```{r}
survey <- read_csv("data/survey_data.csv")
```

## Individuals

hist of frogs, look at means of frog as well
```{r}
survey <- survey %>%
  mutate(frogs = as.numeric(frogs),
         species = as.numeric(species))

# filter out crazy outliers
survey_frogs <- survey %>%
  filter(!is.na(frogs),
         !frogs > 5000)

# need to get mean and median too!
mean(survey_frogs$frogs) # mean of nonoutliers is 1546
median(survey_frogs$frogs) # median of nonoutliers is 1250

# make a histogram
survey_frogs %>%
  ggplot(aes(frogs)) +
  geom_histogram(binwidth = 100)

# geom_freqpoly, I like these better
survey_frogs %>%
  ggplot(aes(frogs)) +
  geom_freqpoly()
```


publishable histogram for ind w/ only entries less than 5000
insert vertical line at mean or median of guesses
insert vertical line at current progress
*updated for final count - final!*
```{r ind_survey_progress_final}
survey_frogs  %>%
  ggplot(aes(frogs)) +
  geom_freqpoly(bins = 50) +
  geom_vline(xintercept = median(survey_frogs$frogs), linetype="solid", 
                color = "red", size=1) +
  geom_vline(xintercept = length(new$Numero), linetype="solid", 
                color = "blue", size=1) +
  labs(x = 'Number of Frogs', y = 'Number of Guesses') +
  annotate(geom="text", x = 3700, y = 14, 
           label = "The median survey guess was 1,200 frogs.",
              color = "red", size = 5) +
  annotate(geom="text", x = 3700, y = 12, 
           label = "We found 1,839!",
              color = "blue", size = 5) +
  theme_classic()
```

## Species

hist of species, look at means and medians
```{r}
survey_species <- survey %>%
  filter(!is.na(species),
    !species > 625)

# mean and medians
mean(survey_species$species)
median(survey_species$species)

# make a histogram
survey_species %>%
  ggplot(aes(species)) +
  geom_histogram(bindwidth = 25)

# geom_freqpoly, I like these better
survey_species %>%
  ggplot(aes(species)) +
  geom_freqpoly()
```

publishable histogram for sp w/ only entries less than 100
insert vertical line at mean or median of guesses
insert vertical line at current progress
*updated for final count, not including genetic sequencing*
```{r sp_survey_progress_final}
survey_species  %>%
  filter(!species > 200) %>%
  ggplot(aes(species)) +
  geom_freqpoly(bins = 50) +
  geom_vline(xintercept = median(survey_species$species), linetype="solid", 
                color = "red", size=1) +
  geom_vline(xintercept = 31, linetype="solid", 
                color = "blue", size=1) +
  labs(x = 'Number of Species', y = 'Number of Guesses') +
  annotate(geom="text", x = 150, y = 10, 
           label = "The median survey guess was 40 species.",
              color = "red", size = 5) +
  annotate(geom="text", x = 150, y = 8.5, 
           label = "Before genetic sequencing,",
              color = "blue", size = 5) +
  annotate(geom="text", x = 150, y = 7.5, 
           label = "we've already found 31!",
              color = "blue", size = 5) +
  theme_classic()
```


--------------------------------------------------------------------------------
# Filtering out list of swabs for David's group to process in January
- 600 total, 300 each season, focus on Priach
- could go for 10 from each of 32 sites, although doubt will have enough

change Final ID to species - easier to work with
```{r}
new <- new %>%
  rename(species = 'Final ID')

# filter out everything not priach
new_priach <- new %>%
  filter(species == 'priach')

# filter by year - start with 2022
new_priach_2022 <- new_priach %>% 
  filter(grepl('/22', Fecha))

# 2023
new_priach_2023 <- new_priach %>% 
  filter(grepl('/23', Fecha))
```

barplot - look at number of priach per site per year
```{r}
# 2022
ggplot(new_priach_2022, aes(Sitio)) +
  geom_bar()

#same thing for 2023:
ggplot(new_priach_2023, aes(Sitio)) +
  geom_bar()
```

Barplot interpretation
2022
- 5 sites don't have 10+ priach in 2022: DB-C, GP-N, JS-N, JZ-N, VC-C
- so 27 sites total (32-5) x 10 = 270

2023
- 8 sites don't have 10+ priach in 2023: 
    - AL-N, DB-C, FLL-B, JS-N, JZJ-C, VC-C, AZ-N, AZ-C
- IMPORTANT - no permission for AZ-N, AZ-C in 2023, so had 0 in 2023
- they got removed from new_priach_2023
- so 24 sites total (32-6-2) x 10 = 240

- 3 sites that don't have 10+ in both - no priach swabs: DB-C, JS-N, VC-C
- 7 sites don't have 10+ in one season:
  - GP-N, JZ-N, AL-N, FLL-B, JZJ-C, AZ-N, AZ-C

- Means 22 sites will have 10 from each season, compare inter-season (440)
- 7 sites from only one season, contribute to intra-season (70)
- fun mix of other species will be second priority (90)

## Pull samples for 1st tier
```{r}
# only sites w/ >9 samples in a season will be included
tierone_2022 <- new_priach_2022 %>%
  group_by(Sitio) %>%
  filter(n() > 9) #%>% # to see how many ind are in 
  #tally()

# create tier one swab list by randomly sampling 10 from this tier one sites
tierone_2022 <- tierone_2022 %>% 
    group_by(Sitio) %>%
    slice_sample(n = 10)
# 270 in tier one from 2022

# same for 2023
tierone_2023 <- new_priach_2023 %>%
  group_by(Sitio) %>%
  filter(n() > 9) #%>%
     #tally()

tierone_2023 <- tierone_2023 %>% 
    group_by(Sitio) %>%
    slice_sample(n = 10)
# 240 in tier one from 2023

# join and then output write.csv these
tierone <- bind_rows(tierone_2022, tierone_2023)
# 510 in this 

write_csv(tierone, "tierone.csv")
## IMPORTANT - b/c of slice sample, exact #s will change - adding _final to official version
```

## Pull samples for 2nd tier
- This will be a mix from 3 different species w/ at least 30 ind
```{r}
# these are all the species with > 30 swabs, regardless of year / site
new_diverse_30 <- new %>%
  filter(species != 'priach') %>%
  group_by(species) %>%
  filter(n() > 30)

new_diverse_30 %>%
  tally()

# randomly sample 30 from these species : barpul, boapel, epibou
tiertwo <- new_diverse_30 %>% 
    group_by(species) %>%
    slice_sample(n = 30)

write_csv(tiertwo, "tiertwo.csv")
## IMPORTANT - b/c of slice sample, exact #s will change - adding _final to official version
```




