
# Comparing diversity, abundance and composition across landuse types

This script will take processed data as input, run statistic tests and produce
figures comparing the dependent variables (diversity, abundance, composition) of
each landuse type broadly. Other scripts will explore the variation in these
response variables w/in specific landuse types.


--------------------------------------------------------------------------------
load packages
```{r}
library(labdsv)
library(tidyverse)
library(vegan)
library(RColorBrewer)
```

load data
## WHY IS 1st COLUMN SHOWING?
```{r}
ranas <- read_csv("data/processed_data/ranas_processed.csv")

new <- read_csv("data/processed_data/new_processed.csv")
```


--------------------------------------------------------------------------------
# Simple richness by type

richness by type - just number of species

## doesn't appear to be any significant differences across landuse types
- although B, V are higher!
```{r}
# group by site and type to retain type for ggplot
richness_by_site <- ranas %>%
  group_by(Sitio, Tipo) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# Group by land_use and calculate species richness
richness_summary <- richness_by_site %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# Create a boxplot
ggplot(richness_by_site, aes(x = Tipo, y = no_species)) +
  geom_boxplot() +
  labs(x = "Land Use Type", y = "Species Richness") +
  ggtitle("Species Richness Across Land Use Types")
```






