
# PROCESSING DATA

Use this script to perform small tasks on raw data to make it tidier, 
easier to work with. Output is processed data to be used in analysis scripts.


# load packages
```{r}
library(tidyverse)
library(vegan)
library(labdsv)
```

define file path for processed outputs
```{r}
# this will be consistent for each output in this script
data_dir <- file.path(".", "data/processed_data")
```


# ranas

ranas - individual occurrence data across all sites
```{r}
# read raw ranas data
ranas <- read_csv("data/raw_data/ranas_04_05_24.csv")

# format dates
ranas$sample_date <- as.Date(ranas$Fecha, format = "%d/%m/%y")

# need to change VC-V to VC-N, turns out this site was not actually abandoned!
ranas <- ranas %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"),
         year = substr(sample_date, 1, 4))

# remove '...20', '...21' column - not sure why these are being added
ranas <- ranas %>%
  select(-c("...20", "...21"))

ranas$Tipo[ranas$Sitio == "VC-N"] <- "N"

# remove any Final_ID with sequence - these are awaiting sequencing results
ranas <- ranas %>% 
  filter(Final_ID != "sequence")

# change all epibou to epiesp 
# Lopez-Hervas et al. 2024 show that epibou in this area is actually epiesp
ranas$Final_ID[ranas$Final_ID == "epibou"] <- "epiesp"

# make processed csv output
filename <- "ranas_processed.csv"
filepath <- file.path(data_dir, filename)

# write df to csv
write_csv(ranas, file = filepath)
```


# new

new - only new individuals from ranas dataset
```{r}
# make new with only new ind
new <- ranas %>%
  filter(Recap == "N")

# # output
# filename <- "new_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(new, file = filepath)
```


# new_no_priach & ranas_no_priach

new_no_priach - remove all priach to look at underlying diversity + abundance
```{r}
# make new with only non-priach ind
new_no_priach <- new %>%
  filter(Final_ID != "priach")

# # output
# filename <- "new_no_priach_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(new_no_priach, file = filepath)

# og ranas with only non-priach ind
ranas_no_priach <- ranas %>%
  filter(Final_ID != "priach")

# # output
# filename <- "ranas_no_priach_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(ranas_no_priach, file = filepath)
```


# new_threatened

new_threatened - only new individuals from sp of conservation concern - 
- listed as DD or higher
```{r}
new_threatened <- new %>%
  filter(Final_ID == 'epiesp' |
         Final_ID == 'lepper' |
         Final_ID == 'oopsyl' |
         Final_ID == 'leubil' |
         Final_ID == 'prilatic'
           )

# # output
# filename <- "new_threatened_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(new_threatened, file = filepath)
```


# new_mat 

matrix for iNEXT package 
- to extrapolate and control for uneven sampling coverage across site types
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat <- new_mat[new_order, ]

# # output
# filename <- "new_mat_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(new_mat, file = filepath)
```


# new_mat_type

this matrix will be at type-level, not site-level, allow for comparisons across
bosque, abandoned, ccn, nacional
```{r}
three_column_type <- new %>%
  group_by(Tipo) %>%
  count(`Final_ID`)

three_column_type <- data.frame(three_column_type)

new_mat_type <- matrify(three_column_type)

# output
# filename <- "new_mat_type_processed.csv"
# filepath <- file.path(data_dir, filename)
# write.csv(new_mat_type, file = filepath, row.names = TRUE)
```


# four new_mat matrices for each site type
- talking with Francisco, seems will need one for each type of site
- independently create sampling coverage curves for each of these
```{r}
# forest
three_column_forest <- new %>%
  filter(Tipo == 'B') %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

three_column_forest <- data.frame(three_column_forest)

new_mat_forest <- matrify(three_column_forest)

# filename <- "new_mat_forest_processed.csv"
# filepath <- file.path(data_dir, filename)
# write.csv(new_mat_forest, file = filepath, row.names = TRUE)

# abandoned
three_column_abandon <- new %>%
  filter(Tipo == 'V') %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

three_column_abandon <- data.frame(three_column_abandon)

new_mat_abandon <- matrify(three_column_abandon)

# filename <- "new_mat_abandon_processed.csv"
# filepath <- file.path(data_dir, filename)
# write.csv(new_mat_abandon, file = filepath, row.names = TRUE)

# shade
three_column_shade <- new %>%
  filter(Tipo == 'N') %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

three_column_shade <- data.frame(three_column_shade)

new_mat_shade <- matrify(three_column_shade)

# filename <- "new_mat_shade_processed.csv"
# filepath <- file.path(data_dir, filename)
# write.csv(new_mat_shade, file = filepath, row.names = TRUE)

# sun
three_column_sun <- new %>%
  filter(Tipo == 'C') %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

three_column_sun <- data.frame(three_column_sun)

new_mat_sun <- matrify(three_column_sun)

# filename <- "new_mat_sun_processed.csv"
# filepath <- file.path(data_dir, filename)
# write.csv(new_mat_sun, file = filepath, row.names = TRUE)
```


# environmental

environmental - environmental data for each sampling event
```{r}
# environmental data - change VC-V to VC-N
environmental <- read_csv("data/raw_data/environment_07_12_24.csv")

environmental$sample_date <- as.Date(environmental$Fecha, format = "%d/%m/%y")

# update VC-N
environmental <- environmental %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"))

environmental$Tipo[environmental$Sitio == "VC-N"] <- "N"

# # output
# filename <- "environmental_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(environmental, file = filepath)
```


# site_data

site_data - location, canopy cover, and miscellaneous data for each site
```{r}
# site_data - change VC-V to VC-N
site_data <- read_csv("data/raw_data/site_data_08_19_24.csv")

# update VC-N
site_data <- site_data %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"))

site_data$Tipo[site_data$Sitio == "VC-N"] <- "N"

# # output
# filename <- "site_data_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(site_data, file = filepath)
```

# local_by_trans

For local - not at a quadrant level, but at a transect level
- Lots of local variables were measured at the quadrant level for practicality

So, here summarize (using mean, coefficient of variation)
- canopy cover
- leaf litter coverage
- vegetation
at a transect level:
```{r}
local_by_trans <- site_data %>%
  rowwise() %>%
  mutate(
    tran_mean_CC = mean(c_across(`CC A`:`CC D`), na.rm = TRUE),
    tran_coef_var_CC = sd(c_across(`CC A`:`CC D`), na.rm = TRUE) / tran_mean_CC
  )
```

Need to fill in vegetation and hojarasca values
- measured during first site visit, if site did not change between subs. visits 
  wrote 'same' - this was the case vast majority of time!
  
So now need to fill these columns based on the values of previous visits:
```{r}
environment_fill <- environmental %>%
  group_by(Sitio, Fecha, Transecto, Bloque) %>%
  mutate(Veg_fill = if_else(Veg == 'same', NA_character_, Veg),
         Hoja_fill = if_else(Hoja =='same', NA_character_, Hoja))

environment_fill <- environment_fill %>%
  group_by(Sitio, Transecto, Bloque) %>%
  fill(Veg_fill, .direction = "down") %>%
  fill(Hoja_fill, .direction = "down") %>%
  ungroup()

print(environment_fill)
```

now can calculate transect visit averages for vegetation and hojarasca
```{r}
environment_fill$Veg_fill <- as.numeric(environment_fill$Veg_fill)
environment_fill$Hoja_fill <- as.numeric(environment_fill$Hoja_fill)

environment_fill_by_trans <- environment_fill %>%
  group_by(Sitio, Fecha, Transecto) %>%
  summarize(tran_mean_veg = mean(Veg_fill, na.rm = TRUE),
            tran_mean_hoja = mean(Hoja_fill, na.rm = TRUE)
            )
```

now join this dataframe with local_by_trans
- ONLY INCLUDING PLANTATIONS HERE! SO ONLY N, C
```{r}
local_by_trans <- left_join(environment_fill_by_trans, local_by_trans,
                                 by = c("Sitio", "Transecto"))

local_by_trans <- local_by_trans %>%
  select(Sitio, Tipo, Fecha, Transecto, 
         tran_mean_veg, tran_mean_hoja, Pair,
         tran_mean_CC, tran_coef_var_CC,
         forest_cover_400, water_dist) %>%
  filter(Tipo %in% c("N", "C"))  # here where filtering out B, V

# bin all transect measurements to categories used in Islas studies

# Define the breaks (bins) and labels
breaks <- c(0, 1, 5, 10, 25, 50, 75, 95, 100)
labels <- c("0", "1-5", "5-10", "10-25", "25-50", "50-75", "75-95", "95-100")

local_by_trans$veg_bins <- cut(local_by_trans$tran_mean_veg, 
                                   breaks = breaks, 
                                   labels = labels, 
                                   include.lowest = TRUE, 
                                   right = FALSE)

local_by_trans$hoja_bins <- cut(local_by_trans$tran_mean_hoja, 
                                   breaks = breaks, 
                                   labels = labels, 
                                   include.lowest = TRUE, 
                                   right = FALSE)

local_by_trans$CC_bins <- cut(local_by_trans$tran_mean_CC, 
                                   breaks = breaks, 
                                   labels = labels, 
                                   include.lowest = TRUE, 
                                   right = FALSE)
```

standardize all variables for local_by_trans for multivariate models
```{r}
local_by_trans <- local_by_trans %>%
  ungroup() %>% # Ungrouping to scale variables
  mutate(tran_mean_veg_scaled = as.numeric(scale(tran_mean_veg)),
         tran_mean_hoja_scaled = as.numeric(scale(tran_mean_hoja)),
         tran_mean_CC_scaled = as.numeric(scale(tran_mean_CC)),
         forest_cover_400_scaled = as.numeric(scale(forest_cover_400)),
         water_dist_scaled = as.numeric(scale(water_dist)))

# filename <- "local_by_trans_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(local_by_trans, file = filepath)
```

# local_by_site
easier to work at site level - removes lots of strange use cases
- use local_by_trans df from above to calculate site-level local variables 
- for each season!
```{r}
# format dates, create year column
local_by_trans$sample_date <- as.Date(local_by_trans$Fecha, format = "%d/%m/%y")
local_by_trans <- local_by_trans %>%
  mutate(year = substr(sample_date, 1, 4))

# calculate year, site-level averages for local vars
local_by_site <- local_by_trans %>%
  group_by(Sitio, year) %>%
  mutate(site_mean_veg = mean(tran_mean_veg),
         site_mean_hoja = mean(tran_mean_hoja),
         site_mean_CC = mean(tran_mean_CC)
  )

# scale all vars
local_by_site <- local_by_site %>%
  ungroup() %>% # Ungrouping to scale variables
  mutate(site_mean_veg_scaled = as.numeric(scale(site_mean_veg)),
         site_mean_hoja_scaled = as.numeric(scale(site_mean_hoja)),
         site_mean_CC_scaled = as.numeric(scale(site_mean_CC)),
         forest_cover_400_scaled = as.numeric(scale(forest_cover_400)),
         water_dist_scaled = as.numeric(scale(water_dist)))
```

# richness_by_site_year
- different than richness by site b/c calculates by year too
```{r}
richness_by_site_year <- new %>%
  group_by(Sitio, year) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))
```

# site_richnes_local
- left_join richness_by_site_year and local_by_site
```{r}
site_richness_local <- left_join(richness_by_site_year, local_by_site,
                                 by = c("Sitio", "year")) %>%
                       filter(Tipo %in% c("N", "C")) %>% # only N,C
                       group_by(Sitio, year) %>% 
                       slice(1) # just take one row per site

# incorporate abundance at site level here
site_abundance <- ranas %>%
  group_by(Sitio, year) %>%
  summarize(num_ind = n_distinct(Numero))

# add to above df to make downstream model creation easier
site_richness_local <- left_join(site_abundance, site_richness_local,
                                 by = c("Sitio", "year")) %>%
                       filter(Tipo %in% c("N", "C"))

# filename <- "site_richness_local_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(site_richness_local, file = filepath)
```

# trans_richness_local
```{r}
trans_richness <- ranas %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_species = n_distinct(Final_ID))

# as currently constructed, this sums richness for each transect on each date 
trans_richness_local <- left_join(trans_richness, local_by_trans,
                                 by = c("Sitio", "Transecto", "Fecha"))

trans_abundance <- ranas %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_ind = n_distinct(Numero))

# add to above df to make downstream model creation easier
trans_richness_local <- left_join(trans_abundance, trans_richness_local,
                                 by = c("Sitio", "Transecto", "Fecha")) %>%
  filter(Tipo %in% c("N", "C"))  # here where filtering out B, V

# filename <- "trans_richness_local_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(trans_richness_local, file = filepath)
```


# landscape_by_site

For landscape - not at a transect level, but at a site level
- Landscape variables were measured at site level for practicality

So, here summarize
- percent surrounding forest cover
- proximity to water
- elevation
at a site level:

# site_type

site_type - retains elevation, tipo data
```{r}
site_type <- site_data %>% # create site_type df - include name, elevation
  filter(Transecto == 1) %>% # so just one row per site
  select(Sitio, Tipo, Elevation, Pair)

# reorder factor levels now for downstream plotting
site_type$Tipo <- factor(site_type$Tipo, c("N", "C", "V", "B"))

# make Pair a factor for glms
# MAY NEED TO SET PAIR AS FACTOR
#site_type$Pair <- factor(site_type$Pair)

# here add harvested column for later analyses
site_type <- site_type %>%
  mutate(harvested = case_when(
    Tipo == "N" ~ "Yes",
    Tipo == "C" ~ "Yes",
    Tipo == "B" ~ "No",
    Tipo == "V" ~ "No")
    )

# # output
# filename <- "site_type_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(site_type, file = filepath)
```


# richness_by_site

richness by site - just number of species
```{r}
# group by site
richness_by_site <- new %>%
  group_by(Sitio) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# join with site_type to carry over tipo, elevation, harvested info
richness_by_site <- richness_by_site %>%
  left_join(site_type, by = "Sitio")

# Group by land_use and calculate species richness
richness_summary <- richness_by_site %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# # output
# filename <- "richness_by_site_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(richness_by_site, file = filepath)
```


# richness_by_site_no_priach

richness_by_site_no_priach - just number of species w/out priach
```{r}
# group by site
richness_by_site_no_priach <- new_no_priach %>%
  group_by(Sitio) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# join with site_type to carry over tipo, elevation, harvested info
richness_by_site_no_priach <- richness_by_site_no_priach %>%
  left_join(site_type, by = "Sitio")

# Group by land_use and calculate species richness
richness_summary_no_priach <- richness_by_site_no_priach %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# # output
# filename <- "richness_by_site_no_priach_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(richness_by_site_no_priach, file = filepath)
```







# FIX WHY IS THIS NOT WORKING? NAS REMAIN WHEN I NEED THEM TO BE ZEROS

# threatened_richness_by_site
number of threatened species
```{r}
# group by site
threatened_richness_by_site <- new_threatened %>%
  group_by(Sitio) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# left_join with site_type to get type, elevation data
threatened_richness_by_site <- site_type %>%
  left_join(threatened_richness_by_site, by = "Sitio")

# change NAs to 0 for plotting
threatened_richness_by_site[is.na(threatened_richness_by_site)] <- 0

# Group by land_use and calculate species richness
threatened_richness_summary <- threatened_richness_by_site %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# # output
# filename <- "threatened_richness_by_site_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(threatened_richness_by_site, file = filepath)
```


# abundance_by_site

just number of ind
```{r}
# note this is same as site_abundance above
abundance_by_site <- new %>%
  group_by(Sitio) %>%
  count()

# join with site_type to carry over tipo, elevation, harvested info
abundance_by_site <- abundance_by_site %>%
  left_join(site_type, by = "Sitio")

abundance_by_site$Tipo <- factor(abundance_by_site$Tipo, c("N", "C", "V", "B"))

# # output
# filename <- "abundance_by_site_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(abundance_by_site, file = filepath)
```


# abundance_by_site_no_priach

just number of non priach ind
```{r}
# note this is same as site_abundance above
abundance_by_site_no_priach <- new_no_priach %>%
  group_by(Sitio) %>%
  count()

# join with site_type to carry over tipo, elevation, harvested info
abundance_by_site_no_priach <- abundance_by_site_no_priach %>%
  left_join(site_type, by = "Sitio")

abundance_by_site_no_priach$Tipo <- factor(abundance_by_site_no_priach$Tipo, 
                                           c("N", "C", "V", "B"))

# # output
# filename <- "abundance_by_site_no_priach_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(abundance_by_site_no_priach, file = filepath)
```


# diversity_by_site

use 3 column format to get diversity indices for each site using vegan package

make new_mat - intermediate step for diversity indices from vegan
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat <- new_mat[new_order, ]
```

diversity_by_site - each major diversity metric
```{r}
# use diversity function from vegan package
diversity_by_site <- site_type %>%
    mutate(Shannon_Index = diversity(new_mat, index = "shannon")) %>%
    mutate(Simpson_Index = diversity(new_mat, index = "simpson")) %>%
    mutate(Inv_Simpson_Index = diversity(new_mat, index = "invsimpson"))

# # output
# filename <- "diversity_by_site_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(diversity_by_site, file = filepath)
```


# diversity_by_site_no_priach

same as above except using new_no_priach
```{r}
three_column_no_priach <- new_no_priach %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column_no_priach <- data.frame(three_column_no_priach)

new_mat_no_priach <- matrify(three_column_no_priach)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat_no_priach <- new_mat_no_priach[new_order, ]
```

diversity_by_site_no_priach - each major diversity metric
```{r}
# use diversity function from vegan package
diversity_by_site_no_priach <- site_type %>%
    mutate(Shannon_Index = diversity(new_mat_no_priach, index = "shannon")) %>%
    mutate(Simpson_Index = diversity(new_mat_no_priach, index = "simpson")) %>%
    mutate(Inv_Simpson_Index = diversity(new_mat_no_priach, index = "invsimpson"))

# # output
# filename <- "diversity_by_site_no_priach_processed.csv"
# filepath <- file.path(data_dir, filename)
# write_csv(diversity_by_site_no_priach, file = filepath)
```










