---
title: "ranas_local_landscape"
author: "Holden Jones"
date: "2024-07-01"
output: html_document
---

# Build and assess models to understand importance of local and landscape vars

```{r}
library(tidyverse)
```

GLM for effect of transect level, local variables on response variables
- species richness, diversity, community composition, abundance
- Spencer suggested using an occupancy approach - also what Dan Karp suggested!

```{r}
ranas <- read_csv("data/processed_data/ranas_processed.csv")
ranas <- as_tibble(ranas)

local_by_trans <- read_csv("data/processed_data/local_by_trans_processed.csv")
local_by_trans <- as_tibble(local_by_trans)

new <- read_csv("data/processed_data/new_processed.csv")
new <- as_tibble(new)

ranas_no_priach <- read_csv("data/processed_data/ranas_no_priach_processed.csv")
ranas_no_priach <- as_tibble(ranas_no_priach)

site_data <- read_csv("data/processed_data/site_data_processed.csv")
site_data <- as_tibble(site_data)
```

# tally sp. & ind. in each transect visit
```{r}
trans_richness <- ranas %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_species = n_distinct(Final_ID))

trans_richness_local <- left_join(trans_richness, local_by_trans,
                                 by = c("Sitio", "Transecto", "Fecha"))

trans_abundance <- ranas %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_ind = n_distinct(Numero))

# add to above df to make downstream model creation easier
trans_richness_local <- left_join(trans_abundance, trans_richness_local,
                                 by = c("Sitio", "Transecto", "Fecha")) %>%
  filter(Tipo %in% c("N", "C"))  # here where filtering out B, V
```

# non priach: tally sp. & ind. in each transect visit 
```{r}
trans_richness_no_priach <- ranas_no_priach %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_species = n_distinct(Final_ID))

trans_richness_local_no_priach <- left_join(trans_richness_no_priach, 
                                            local_by_trans,
                                 by = c("Sitio", "Transecto", "Fecha"))

trans_abundance_no_priach <- ranas_no_priach %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_ind = n_distinct(Numero))


### THIS MIGHT NOT WORK ###
# add to above df to make downstream model creation easier
trans_richness_local_no_priach <- left_join(trans_abundance_no_priach, 
                                            trans_richness_local_no_priach,
                                 by = c("Sitio", "Transecto", "Fecha")) %>%
  filter(Tipo %in% c("N", "C"))  # here where filtering out B, V
```

throw together some glms! starting point!
# don't have a random effect here - probably would want to use this? 
- can't run random effects w/ glm: 1 | Sitio
```{r}
model_richness_trans <- glm(num_species ~ 
                              tran_mean_veg + 
                              tran_mean_hoja + 
                              tran_mean_CC,
                            data = trans_richness_local, 
                            family = poisson()
                            )

# Display the model summary
summary(model_richness_trans)
```

# quick figures here! just for first draft of MS
```{r}
# tran_mean_hoja on num_species
ggplot(trans_richness_local, aes(tran_mean_hoja, num_species)) + 
       geom_point() +
  geom_smooth()

# tran_mean_CC on num_species
ggplot(trans_richness_local, aes(tran_mean_CC, num_species)) + 
       geom_point() +
  geom_smooth()

# tran_mean_veg on num_species
ggplot(trans_richness_local, aes(tran_mean_veg, num_species)) + 
       geom_point() +
  geom_smooth()
```

# Single glms for local var with species richness

# richness ~ veg
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_veg <- glm(num_species ~ tran_mean_veg, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_rich_veg)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_veg = seq(min(trans_richness_local$tran_mean_veg),
                      max(trans_richness_local$tran_mean_veg), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_veg, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = tran_mean_veg, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Species vs. Mean Vegetation",
       x = "Mean Vegetation Cover",
       y = "Number of Species")
```

# richness ~ hoja
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_hoja <- glm(num_species ~ tran_mean_hoja, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_rich_hoja)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_hoja = seq(min(trans_richness_local$tran_mean_hoja),
                      max(trans_richness_local$tran_mean_hoja), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_hoja, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = tran_mean_hoja, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Species vs. Mean Leaf Cover",
       x = "Mean Leaf Cover",
       y = "Number of Species")
```

# richness ~ CC
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_cc <- glm(num_species ~ tran_mean_CC, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_rich_cc)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_CC = seq(min(trans_richness_local$tran_mean_CC),
                      max(trans_richness_local$tran_mean_CC), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_cc, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = tran_mean_CC, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Species vs. Mean Canopy Cover",
       x = "Mean Canopy Cover Cover",
       y = "Number of Species")
```

# richness ~ forest cover
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_forest <- glm(num_species ~ forest_cover_400, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_rich_forest)

# Create a new data frame for predictions
new_data <- data.frame(
  forest_cover_400 = seq(min(trans_richness_local$forest_cover_400),
                      max(trans_richness_local$forest_cover_400), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_forest, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = forest_cover_400, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Species vs. Percent Surrounding Forest Cover",
       x = "Percent Surrounding Forest Cover (400m)",
       y = "Number of Species")
```

# richness ~ freshwater distance
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_water <- glm(num_species ~ water_dist, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_rich_water)

# Create a new data frame for predictions
new_data <- data.frame(
  water_dist = seq(min(trans_richness_local$water_dist),
                      max(trans_richness_local$water_dist), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_water, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = water_dist, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Species vs. Distance to Freshwater",
       x = "Distance to Freshwater (m)",
       y = "Number of Species")
```



# Single glms for local var with abundance

# abundance ~ veg
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_veg <- glm(num_ind ~ tran_mean_veg, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_abun_veg)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_veg = seq(min(trans_richness_local$tran_mean_veg),
                      max(trans_richness_local$tran_mean_veg), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_abun <- predict(
                                 poisson_abun_veg, 
                                 newdata = new_data, 
                                 type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = tran_mean_veg, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Individuals vs. Mean Vegetation",
       x = "Mean Vegetation Cover",
       y = "Number of Individuals")
```

# abundance ~ hoja
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_hoja <- glm(num_ind ~ tran_mean_hoja, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_abun_hoja)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_hoja = seq(min(trans_richness_local$tran_mean_hoja),
                      max(trans_richness_local$tran_mean_hoja), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_abun <- predict(
                                 poisson_abun_hoja, 
                                 newdata = new_data, 
                                 type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = tran_mean_hoja, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Individuals vs. Mean Leaf Coverage",
       x = "Mean Leaf Cover",
       y = "Number of Individuals")
```

# abundance ~ CC
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_CC <- glm(num_ind ~ tran_mean_CC, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_abun_CC)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_CC = seq(min(trans_richness_local$tran_mean_CC),
                      max(trans_richness_local$tran_mean_CC), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_abun_CC, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = tran_mean_CC, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Individuals vs. Mean Canopy Coverage",
       x = "Mean Canopy Cover",
       y = "Number of Individuals")
```


# abundance ~ forest cover
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_forest <- glm(num_ind ~ forest_cover_400, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_abun_forest)

# Create a new data frame for predictions
new_data <- data.frame(
  forest_cover_400 = seq(min(trans_richness_local$forest_cover_400),
                      max(trans_richness_local$forest_cover_400), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_abun_forest, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = forest_cover_400, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Individuals vs. Percent Forest Cover",
       x = "Percent Forest Cover Within 400 m",
       y = "Number of Individuals")
```


# abundance ~ freshwater distance
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_water <- glm(num_ind ~ water_dist, 
                        data = trans_richness_local, 
                        family = poisson())

summary(poisson_abun_water)

# Create a new data frame for predictions
new_data <- data.frame(
  water_dist = seq(min(trans_richness_local$water_dist),
                      max(trans_richness_local$water_dist), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_abun_water, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local, 
       aes(x = water_dist, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Individuals vs. Proximity to Fresh Water",
       x = "Distance to Fresh Water (m)",
       y = "Number of Individuals")
```





# Single glms for local var with non_priach species richness

# non_priach_richness ~ veg
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_veg_no_priach <- glm(num_species ~ tran_mean_veg, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_rich_veg_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_veg = seq(min(trans_richness_local_no_priach$tran_mean_veg),
                      max(trans_richness_local_no_priach$tran_mean_veg), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_veg_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = tran_mean_veg, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus Species vs. Mean Vegetation",
       x = "Mean Vegetation Cover",
       y = "Number of Species")
```

# non_priach_richness ~ hoja
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_hoja_no_priach <- glm(num_species ~ tran_mean_hoja, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_rich_hoja_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_hoja = seq(min(trans_richness_local_no_priach$tran_mean_hoja),
                      max(trans_richness_local_no_priach$tran_mean_hoja), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_hoja_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = tran_mean_hoja, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus species vs. Mean Leaf Cover",
       x = "Mean Leaf Cover",
       y = "Number of Species")
```

# non_priach_richness ~ CC
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_CC_no_priach <- glm(num_species ~ tran_mean_CC, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_rich_CC_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_CC = seq(min(trans_richness_local_no_priach$tran_mean_CC),
                      max(trans_richness_local_no_priach$tran_mean_CC), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_CC_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = tran_mean_CC, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus species vs. Mean Canopy Cover",
       x = "Mean Leaf Cover",
       y = "Number of Species")
```


# non_priach_richness ~ forest cover
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_forest_no_priach <- glm(num_species ~ forest_cover_400, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_rich_forest_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  forest_cover_400 = seq(min(trans_richness_local_no_priach$forest_cover_400),
                      max(trans_richness_local_no_priach$forest_cover_400), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_forest_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = forest_cover_400, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus species vs. Percent Forest Cover",
       x = "Percent Forest Cover Within 400 m",
       y = "Number of Species")
```


# non_priach_richness ~ freshwater distance
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_rich_water_no_priach <- glm(num_species ~ water_dist, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_rich_water_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  water_dist = seq(min(trans_richness_local_no_priach$water_dist),
                      max(trans_richness_local_no_priach$water_dist), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_rich_water_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = water_dist, y = num_species)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus species vs. Proximity to Fresh Water",
       x = "Proximity to Fresh Water (m)",
       y = "Number of Species")
```



# Single glms for local var with non_priach abundance

# non_priach_abundance ~ veg
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_veg_no_priach <- glm(num_ind ~ tran_mean_veg, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_abun_veg_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_veg = seq(min(trans_richness_local_no_priach$tran_mean_veg),
                      max(trans_richness_local_no_priach$tran_mean_veg), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_abun <- predict(
                                 poisson_abun_veg_no_priach, 
                                 newdata = new_data, 
                                 type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = tran_mean_veg, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus individuals vs. Mean Vegetation",
       x = "Mean Vegetation Cover",
       y = "Number of Individuals")
```

# non_priach_abundance ~ hoja
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_hoja_no_priach <- glm(num_ind ~ tran_mean_hoja, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_abun_hoja_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_hoja = seq(min(trans_richness_local_no_priach$tran_mean_hoja),
                      max(trans_richness_local_no_priach$tran_mean_hoja), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_abun <- predict(
                                 poisson_abun_hoja_no_priach, 
                                 newdata = new_data, 
                                 type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = tran_mean_hoja, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus individuals vs. Mean Leaf Coverage",
       x = "Mean Leaf Cover",
       y = "Number of Individuals")
```

# non_priach_abundance ~ CC
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_CC_no_priach <- glm(num_ind ~ tran_mean_CC, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_abun_CC_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  tran_mean_CC = seq(min(trans_richness_local_no_priach$tran_mean_CC),
                      max(trans_richness_local_no_priach$tran_mean_CC), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_abun_CC_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = tran_mean_CC, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus individuals vs. Mean Canopy Coverage",
       x = "Mean Canopy Cover",
       y = "Number of Individuals")
```


# non_priach_abundance ~ forest cover
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_forest_no_priach <- glm(num_ind ~ forest_cover_400, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_abun_forest_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  forest_cover_400 = seq(min(trans_richness_local_no_priach$forest_cover_400),
                      max(trans_richness_local_no_priach$forest_cover_400), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_abun_forest_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = forest_cover_400, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus individuals vs. Percent Forest Cover",
       x = "Percent Forest Cover Within 400 m",
       y = "Number of Individuals")
```


# non_priach_abundance ~ freshwater distance
- Poisson dist.
- using code from CHAT GPT:
```{r}
# Fit the Poisson GLM
poisson_abun_water_no_priach <- glm(num_ind ~ water_dist, 
                        data = trans_richness_local_no_priach, 
                        family = poisson())

summary(poisson_abun_water_no_priach)

# Create a new data frame for predictions
new_data <- data.frame(
  water_dist = seq(min(trans_richness_local_no_priach$water_dist),
                      max(trans_richness_local_no_priach$water_dist), 
                      length.out = 100))

# Add predicted values to the new data frame
new_data$predicted_species <- predict(
                                      poisson_abun_water_no_priach, 
                                      newdata = new_data, 
                                      type = "response")

# Plot the data and the fitted model
ggplot(trans_richness_local_no_priach, 
       aes(x = water_dist, y = num_ind)) +
  geom_jitter(width = 0, height = 0.1) + 
  geom_smooth() +
  labs(title = "Poisson Regression: Number of Non P. achatinus individuals vs. Proximity to Fresh Water",
       x = "Proximity to Freshwater (m)",
       y = "Number of Individuals")
```







#PALMER PENGUINS

How to run a glmmer with random effects?
## palmer penguins tutorial
- https://www.youtube.com/watch?v=smYZdlbE9m8&t=92s
```{r}
library(tidyverse)
library(palmerpenguins)
library(glmmTMB)
library(DHARMa)
library(ggplot2)
library(scico)
library(Matrix)
```


# Incorporating random effects
- Really we are most interested in landscape level random effects, and veg.
- Perhaps would use forest cover and water proximity as fixed effects, but these
  are not things that would be manipulated as part of management regime (well, forest cover could)
  so maybe best as random effects?
  
# 5 models for transect level species diversity with random effects
```{r}
# species against hoja with veg as random effect
mod1 <- glmmTMB(num_species ~ tran_mean_hoja + (1|tran_mean_veg),
                data = trans_richness_local, 
                        family = poisson())

summary(mod1)
coef(mod1)
# AIC 1143

# species against hoja with canopy cover as random effect
mod2 <- glmmTMB(num_species ~ tran_mean_hoja + (1|tran_mean_CC),
                data = trans_richness_local, 
                        family = poisson())

summary(mod2)
coef(mod2)
# AIC 1143

# species against hoja with no random effect
mod3 <- glmmTMB(num_species ~ tran_mean_hoja,
                data = trans_richness_local, 
                        family = poisson())

summary(mod3)
coef(mod3)
# AIC 1141

# species against hoja with pair as random effect
mod4 <- glmmTMB(num_species ~ tran_mean_hoja + (1|Pair),
                data = trans_richness_local, 
                        family = poisson())

summary(mod4)
coef(mod4)
# AIC 1137

# species against forest cover with pair as random effect
mod5 <- glmmTMB(num_species ~ forest_cover_400 +(1|Pair),
                data = trans_richness_local, 
                        family = poisson())

summary(mod5)
coef(mod5)
# AIC 1138


# species against water_dist with pair as random effect
mod6 <- glmmTMB(num_species ~ water_dist,
                data = trans_richness_local, 
                        family = poisson())

summary(mod6)
coef(mod6)
# AIC 1143
```














