
# Processing raw data

Use this script to perform small tasks on raw data to make it tidier, 
easier to work with. Output is processed data to be used in analysis scripts.

## next step is to update sequencing data


--------------------------------------------------------------------------------
load packages
```{r}
library(tidyverse)
```

define file path for processed outputs
```{r}
# this will be consistent for each output in this script
data_dir <- file.path(".", "data/processed_data")
```


--------------------------------------------------------------------------------
ranas - individual occurrence data across all sites
```{r}
# read raw ranas data
ranas <- read_csv("data/raw_data/ranas_05_17_23.csv")

# format dates
ranas$sample_date <- as.Date(ranas$Fecha, format = "%d/%m/%y")

# need to change VC-V to VC-N, turns out this site was not actually abandoned!
ranas <- ranas %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"),
         year = substr(sample_date, 1, 4))

ranas$Tipo[ranas$Sitio == "VC-N"] <- "N"

# make processed csv output
filename <- "ranas_processed.csv"
filepath <- file.path(data_dir, filename)

# write df to csv
write.csv(ranas, file = filepath)
```

## add in identification data for sequenced individuals - perhaps do by hand?
```{r}

```


--------------------------------------------------------------------------------
new - only new individuals from ranas dataset
```{r}
# make new with only new ind
new <- ranas %>%
  filter(Recap == "N")

# output
filename <- "new_processed.csv"
filepath <- file.path(data_dir, filename)
write.csv(new, file = filepath)
```


--------------------------------------------------------------------------------
environmental - environmental data for each sampling event
```{r}
# environmental data - change VC-V to VC-N
environmental <- read_csv("data/raw_data/environment_04_08_23.csv")

environmental$sample_date <- as.Date(environmental$Fecha, format = "%d/%m/%y")

# update VC-N
environmental <- environmental %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"))

environmental$Tipo[environmental$Sitio == "VC-N"] <- "N"

# output
filename <- "environmental_processed.csv"
filepath <- file.path(data_dir, filename)
write.csv(environmental, file = filepath)
```

--------------------------------------------------------------------------------
site_data - location, canopy cover, and miscellaneous data for each site
```{r}
# site_data - change VC-V to VC-N
site_data <- read_csv("data/raw_data/site_data.csv")

# update VC-N
site_data <- site_data %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"))

site_data$Tipo[site_data$Sitio == "VC-N"] <- "N"

# output
filename <- "site_data_processed.csv"
filepath <- file.path(data_dir, filename)
write.csv(site_data, file = filepath)
```


