---
title: "ranas_analysis"
author: "Holden Jones"
date: "2024-10-23"
output: html_document
---

Use this document for analyses and figures to be used in paper

load packages
```{r, include=FALSE}
library(labdsv)
library(tidyverse)
library(vegan)
library(iNEXT)
library(sjmisc)
library(RColorBrewer) # for assigning plot colors
library(lme4)
library(ggpubr) # for arranging multi-panel figures
library(car)
library(glmmTMB)
library(httr)
library(DHARMa)
library(ggrepel)
```

load data
```{r, include=FALSE}
ranas <- read_csv("data/processed_data/ranas_processed.csv")
ranas <- as_tibble(ranas)

new <- read_csv("data/processed_data/new_processed.csv")
new <- as_tibble(new)

new_no_priach <- read_csv("data/processed_data/new_no_priach_processed.csv")
new_no_priach <- as_tibble(new_no_priach)

new_threatened <- read_csv("data/processed_data/new_threatened_processed.csv")
new_threatened <- as_tibble(new_threatened)

### REVIEW THIS SECTION ###
new_mat <- read_csv("data/processed_data/new_mat_processed.csv")
new_mat <- as.matrix(new_mat)

new_mat_test <- read_csv("data/processed_data/new_mat_test_processed.csv")
new_mat_test <- as.matrix(new_mat_test)
### REVIEW THIS SECTION ###

site_data <- read_csv("data/processed_data/site_data_processed.csv")
site_data <- as_tibble(site_data)

site_type <- read_csv("data/processed_data/site_type_processed.csv")
site_type <- as_tibble(site_type)

richness_by_site <- read_csv(
  "data/processed_data/richness_by_site_processed.csv")
richness_by_site <- as_tibble(richness_by_site)

richness_by_site_no_priach <- read_csv(
  "data/processed_data/richness_by_site_no_priach_processed.csv")
richness_by_site_no_priach <- as_tibble(richness_by_site_no_priach)

threatened_richness_by_site <- read_csv(
  "data/processed_data/threatened_richness_by_site_processed.csv")
threatened_richness_by_site <- as_tibble(threatened_richness_by_site)

abundance_by_site <- read_csv(
  "data/processed_data/abundance_by_site_processed.csv")
abundance_by_site <- as_tibble(abundance_by_site)

abundance_by_site_no_priach <- read_csv(
  "data/processed_data/abundance_by_site_no_priach_processed.csv")
abundance_by_site <- as_tibble(abundance_by_site_no_priach)

diversity_by_site <- read_csv(
  "data/processed_data/diversity_by_site_processed.csv")
diversity_by_site <- as_tibble(diversity_by_site)

diversity_by_site_no_priach <- read_csv(
  "data/processed_data/diversity_by_site_no_priach_processed.csv")
diversity_by_site_no_priach <- as_tibble(diversity_by_site_no_priach)

trans_richness_local <- read_csv(
  "data/processed_data/trans_richness_local_processed.csv")
trans_richness_local <- as_tibble(trans_richness_local)

site_richness_local <- read_csv(
  "data/processed_data/site_richness_local_processed.csv")
site_richness_local <- as_tibble(site_richness_local)

priach_by_site <- read_csv(
  "data/processed_data/priach_by_site_processed.csv")
priach_by_site <- as_tibble(priach_by_site)

# matrix files for each landuse type
data.shade=read.table("data/processed_data/SHADE.txt", head= T)
data.sun=read.table("data/processed_data/SUN.txt", head= T)
data.for=read.table("data/processed_data/FOREST.txt", head= T)
data.aba=read.table("data/processed_data/ABANDON.txt", head= T)

trans_richness_local_no_priach <- read_csv(
  "data/processed_data/trans_richness_local_no_priach_processed.csv")
trans_richness_local_no_priach <- as_tibble(trans_richness_local_no_priach)

site_richness_local_no_priach <- read_csv(
  "data/processed_data/site_richness_local_no_priach_processed.csv")
site_richness_local_no_priach <- as_tibble(site_richness_local_no_priach)

MATFROGH <- read_csv("data/processed_data/MATFROGH.csv")
MATFROGH <- as_tibble(MATFROGH)
```

define a consistent palet, labels, and theme to be used in figures
```{r, include=FALSE}
# palet for comparisons of all 4 site types
pal <- brewer.pal(4, "BrBG")

# set a theme for ggplots
my_theme <- function() {
  theme_minimal() + 
   theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 16, color = "gray25"),
        axis.title = element_text(size = 22, color = "gray25"),
        legend.text = element_text(size = 16))
}
```

functions
```{r}
# function for checking model overdispersion
check_overdispersion <- function(model) {
  # Residual deviance and degrees of freedom
  dev <- deviance(model)
  df <- df.residual(model)
  
  # Overdispersion ratio
  ratio <- dev / df
  
  # Test statistic (chi-square distribution)
  p_value <- pchisq(dev, df, lower.tail = FALSE)
  
  # Display the results
  c(Overdispersion = ratio, p_value = p_value)
}
```


#1. Compare abundance, richness and Shannon diversity of shade and sun

##a. Richness
- Use two-sample t-test b/c not perfectly paired in landscape and dif number of sites
- If remove VC-N (weird site turned out to be N, not paired), then can run paired t-test
```{r}
# create species richness vector for each cacao type
rich_ccn <- richness_by_site %>%
  filter(Tipo == 'C') %>%
  pull(no_species)
  
rich_n <- richness_by_site %>%
  filter(Tipo == 'N') %>%
  pull(no_species)

# run two-sample t-test for species richness
t.test(rich_ccn, rich_n, paired = FALSE)
```

- Here run paired t-test w/out VC-N
```{r}
# drop VC-N from df, this site messes up pairing for t-test
pair_rich_by_site <- richness_by_site %>%
  filter(Sitio != 'VC-N') %>%
  arrange(Pair) # sort by Pair so vectors pulled align with pairs

# create species richness vector for each cacao type
pair_rich_ccn <- pair_rich_by_site %>%
  filter(Tipo == 'C') %>%
  pull(no_species)
  
pair_rich_n <- pair_rich_by_site %>%
  filter(Tipo == 'N') %>%
  pull(no_species)

t.test(pair_rich_ccn, pair_rich_n, paired = TRUE)
```

Boxplot figure for CCN vs. Nacional species richness
```{r}
# site labels for 2 site comparison
site_labels <-  c("Shade \n (n = 12)", "Sun \n (n = 11)")

# reorder factor levels
richness_by_site$Tipo <- factor(richness_by_site$Tipo,
                         c("N", "C", "V", "B"))

richness_by_site_cacao <- richness_by_site %>%
  filter(Tipo == "C" | Tipo == 'N')

# plot species richness across landuse types using palet, my_theme
plot_richness_by_site_cacao <- ggplot(richness_by_site_cacao, 
                                     aes(x = Tipo, y = no_species, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Cacao type",
       y = "Species richness")
plot_richness_by_site_cacao
#ggsave("output/plot_richness_by_site_cacao.png", width = 10, height = 10)

# below using ggarrange to align multiple figs - don't want to repeat axis labels
plot_richness_by_site_cacao_arrange <- ggplot(richness_by_site_cacao, 
                                     aes(x = Tipo, y = no_species, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + 
  scale_y_continuous(limits = c(2, 10)) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  theme(axis.title.x=element_blank(),
        #axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(y = "Richness")
plot_richness_by_site_cacao_arrange
```


##b. Shannon diversity
Compare Shannon diversity of CCN and Nacional
- same process as above, can be done with paired or two-sample t-test

two-sample t-test
```{r}
# Two sample t-test for shannon diversity:

# create Shannon diversity vector for each cacao type
shannon_ccn <- diversity_by_site %>%
  filter(Tipo == 'C') %>%
  pull(Shannon_Index)
  
shannon_n <- diversity_by_site %>%
  filter(Tipo == 'N') %>%
  pull(Shannon_Index)

# run two-sample t-test for shannon diversity
t.test(shannon_ccn, shannon_n, paired = FALSE)
```

paired t-test
```{r}
# Paired t-test for Shannon diversity 

# drop VC-N from df, this site messes up pairing for t-test
pair_diversity_by_site <- diversity_by_site %>%
  filter(Sitio != 'VC-N') %>%
  arrange(Pair) # sort by Pair so vectors pulled align with pairs

# create Shannon diversity vector for each cacao type
pair_shannon_ccn <- pair_diversity_by_site %>%
  filter(Tipo == 'C') %>%
  pull(Shannon_Index)
  
pair_shannon_n <- pair_diversity_by_site %>%
  filter(Tipo == 'N') %>%
  pull(Shannon_Index)

t.test(pair_shannon_ccn, pair_shannon_n, paired = TRUE)
```

Boxplot figure for CCN vs. Nacional Shannon diversity
```{r}
# reorder factor levels
diversity_by_site$Tipo <- factor(diversity_by_site$Tipo,
                         c("N", "C", "V", "B"))

shannon_by_site_cacao <- diversity_by_site %>%
  filter(Tipo == "C" | Tipo == 'N')

# plot Shannon diversity across landuse types using palet, my_theme
plot_shannon_by_site_cacao <- ggplot(shannon_by_site_cacao, 
                                     aes(x = Tipo, y = Shannon_Index, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Cacao type",
       y = "Shannon index")

plot_shannon_by_site_cacao
#ggsave("output/plot_shannon_by_site_cacao.png", width = 10, height = 10)

# below using ggarrange to align multiple figs
plot_shannon_by_site_cacao_arrange <- ggplot(shannon_by_site_cacao, 
                                     aes(x = Tipo, y = Shannon_Index, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(labels = c(site_labels)) +
  scale_y_continuous(limits = c(0, 1.6)) +
  my_theme() + 
  labs(x = "Cacao type",
       y = "Shannon index")

plot_shannon_by_site_cacao_arrange
```



##c. Abundance
Same process as above - use abundance_by_site for two sample t-test
```{r}
# create species abundance vector for each cacao type
abun_ccn <- abundance_by_site %>%
  filter(Tipo == 'C') %>%
  pull(n)
  
abun_n <- abundance_by_site %>%
  filter(Tipo == 'N') %>%
  pull(n)

# run two-sample t-test for species richness
t.test(abun_ccn, abun_n, paired = FALSE)
```

paired t-test
```{r}
# drop VC-N from df, this site messes up pairing for t-test
pair_abundance_by_site <- abundance_by_site %>%
  filter(Sitio != 'VC-N') %>%
  arrange(Pair) # sort by Pair so vectors pulled align with pairs

# create abundance vector for each cacao type
pair_abun_ccn <- pair_abundance_by_site %>%
  filter(Tipo == 'C') %>%
  pull(n)
  
pair_abun_n <- pair_abundance_by_site %>%
  filter(Tipo == 'N') %>%
  pull(n)

t.test(pair_abun_ccn, pair_abun_n, paired = TRUE)
```

Boxplot figure for CCN vs. Nacional abundance
```{r}
# site labels for 2 site comparison
site_labels <-  c("Shade \n (n = 12)", "Sun \n (n = 11)")

# reorder factor levels
abundance_by_site$Tipo <- factor(abundance_by_site$Tipo,
                         c("N", "C", "V", "B"))

abundance_by_site_cacao <- abundance_by_site %>%
  filter(Tipo == "C" | Tipo == 'N')

# plot species richness across landuse types using palet, my_theme
plot_abundance_by_site_cacao <- ggplot(abundance_by_site_cacao, 
                                     aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Cacao type",
       y = "Amphibian abundance")

plot_abundance_by_site_cacao
#ggsave("output/plot_abundance_by_site_cacao.png", width = 10, height = 10)

# below using ggarrange to align multiple figs - don't want to repeat axis labels
plot_abundance_by_site_cacao_arrange <- ggplot(abundance_by_site_cacao, 
                                     aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + 
  scale_y_continuous(limits = c(0, 51)) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  theme(axis.title.x=element_blank(),
        #axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(y = "Abundance")
plot_abundance_by_site_cacao_arrange
```


#2.Compare abundance, richness and Shannon diversity of forest and abandoned plantations

##a. NMDS ordination plot - this should go below with contextualize??!
- using new_mat df, species-by-site abundance matrix
- draw boundaries of landuse types
- possible that this figure doesn't ultimately get included in MS
```{r}
#### Order new_mat df ####
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat <- new_mat[new_order, ]


#### Make ordination plot ####

# assign treatment levels per landuse type on reordered new_mat
landuse_treatments=c(rep("Shade",12),
                      rep("Sun",11), 
                      rep("Abandoned",4), 
                      rep("Forest",5))

# use community by species matrix - stress = 0.26 - okay representation
new_mat_NMDS = metaMDS(new_mat, k = 2)

# sites are open circles, species are red crosses
plot(new_mat_NMDS)


# help visualize by labeling specific sites and species

# ordination plot function especially for congested plots
ordiplot(new_mat_NMDS,type="n")
# this function adds text or points to ordination plots
orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

orditorp(new_mat_NMDS,display="sites",cex=1.25,air=0.01)

ordiplot(new_mat_NMDS,type="n")

ordihull(new_mat_NMDS,groups=landuse_treatments,draw="polygon",
         col="grey90",label=F)

orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

# here colors need to align with landuse_treatments from above!
# of course need to use chocolate!
orditorp(new_mat_NMDS,display="sites",col=c(rep("chocolate4",12),
                                            rep("chocolate1",11),
                                            rep("lightgreen",4),
                                            rep("darkgreen",5)), 
         air=0.01,cex=1.25)
```


##b. Richness
- Use two-sample t-test b/c not paired and dif number of sites
```{r}
# create species richness vector for each cacao type
rich_forest <- richness_by_site %>%
  filter(Tipo == 'B') %>%
  pull(no_species)
  
rich_abandon <- richness_by_site %>%
  filter(Tipo == 'V') %>%
  pull(no_species)

# run two-sample t-test for species richness
t.test(rich_forest, rich_abandon, paired = FALSE)
```

Boxplot figure for forest vs. abandoned species richness
```{r}
site_labels <- c("Abandoned \n (n = 4)", "Forest \n (n = 5)")

# reorder factor levels
richness_by_site$Tipo <- factor(richness_by_site$Tipo,
                         c("N", "C", "V", "B"))

richness_by_site_inactive <- richness_by_site %>%
  filter(Tipo == "B" | Tipo == 'V')

# plot species richness across landuse types using palet, my_theme
plot_richness_by_site_inactive <- ggplot(richness_by_site_inactive, 
                                     aes(x = Tipo, y = no_species, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#80CDC1", "#018571")) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Reference type",
       y = "Species richness")

plot_richness_by_site_inactive
#ggsave("output/plot_richness_by_site_inactive.png", width = 10, height = 10)

# below using ggarrange to align multiple figs - don't want to repeat x axes labels
plot_richness_by_site_inactive_arrange <- ggplot(richness_by_site_inactive, 
                                     aes(x = Tipo, y = no_species, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#80CDC1", "#018571")) + 
  scale_y_continuous(limits = c(2, 10)) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  theme(axis.title.x = element_blank(),
       # axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(), #,
       # axis.text.y = element_blank() ,
        axis.ticks.y = element_blank()
        )
plot_richness_by_site_inactive_arrange
```


##c. Shannon diversity
Compare Shannon diversity of abandoned plantations and remnant forest

two-sample t-test
```{r}
# Two sample t-test for shannon diversity:

# create Shannon diversity vector for each type
shannon_abandon <- diversity_by_site %>%
  filter(Tipo == 'V') %>%
  pull(Shannon_Index)
  
shannon_forest <- diversity_by_site %>%
  filter(Tipo == 'B') %>%
  pull(Shannon_Index)

# run two-sample t-test for shannon diversity
t.test(shannon_abandon, shannon_forest, paired = FALSE)
```

Boxplot figure for abandoned vs. forest Shannon diversity
```{r}
site_labels <- c("Abandoned \n (n = 4)", "Forest \n (n = 5)")

# reorder factor levels
diversity_by_site$Tipo <- factor(diversity_by_site$Tipo,
                         c("N", "C", "V", "B"))

shannon_by_site_inactive <- diversity_by_site %>%
  filter(Tipo == "V" | Tipo == 'B')

# plot Shannon diversity across landuse types using palet, my_theme
plot_shannon_by_site_inactive <- ggplot(shannon_by_site_inactive, 
                                     aes(x = Tipo, y = Shannon_Index, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#80CDC1", "#018571")) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Reference type",
       y = "Shannon diversity")

plot_shannon_by_site_inactive
#ggsave("output/plot_shannon_by_site_inactive.png", width = 10, height = 10)

# below using ggarrange to align multiple figs - here don't want y axis legends
plot_shannon_by_site_inactive_arrange <- ggplot(shannon_by_site_inactive, 
                                     aes(x = Tipo, y = Shannon_Index, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#80CDC1", "#018571")) + 
  scale_x_discrete(labels = c(site_labels)) +
  scale_y_continuous(limits = c(0, 1.6)) +
  my_theme() + 
  labs(x = "Reference type") +
  theme(axis.title.y = element_blank(),
        #axis.text.y = element_blank() #,
        axis.ticks.y = element_blank()
        )
plot_shannon_by_site_inactive_arrange
```


##d. Abundance
Same process as above - use abundance_by_site for two sample t-test
```{r}
# create species abundance vector for each cacao type
abun_abandon <- abundance_by_site %>%
  filter(Tipo == 'V') %>%
  pull(n)
  
abun_forest <- abundance_by_site %>%
  filter(Tipo == 'B') %>%
  pull(n)

# run two-sample t-test for species richness
t.test(abun_abandon, abun_forest, paired = FALSE)
```

Boxplot figure for forest vs. abandoned abundance
```{r}
# reorder factor levels
abundance_by_site$Tipo <- factor(abundance_by_site$Tipo,
                         c("N", "C", "V", "B"))

abundance_by_site_inactive <- abundance_by_site %>%
  filter(Tipo == "V" | Tipo == 'B')

# plot Shannon diversity across landuse types using palet, my_theme
plot_abundance_by_site_inactive <- ggplot(abundance_by_site_inactive, 
                                     aes(x = Tipo, y = n, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#80CDC1", "#018571")) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Reference type",
       y = "Amphibian abundance")
plot_abundance_by_site_inactive
#ggsave("output/plot_abundance_by_site_inactive.png", width = 10, height = 10)

# below using ggarrange to align multiple figs, don't want to repeat axes
plot_abundance_by_site_inactive_arrange <- ggplot(abundance_by_site_inactive, 
                                     aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#80CDC1", "#018571")) + 
  scale_y_continuous(limits = c(0, 51)) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  theme(axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
plot_abundance_by_site_inactive_arrange
```


##e. Arrange figures for publication
arrange these four or six figures as panels in one window
```{r}
# richness and shannon diversity
compare_rich_shannon <- ggarrange(
  plot_richness_by_site_cacao + rremove("xlab"), 
  plot_richness_by_site_inactive + rremove("xlab") + rremove("ylab"),
  plot_shannon_by_site_cacao,
  plot_shannon_by_site_inactive + rremove("ylab"),
  labels = c("A", "B", "C", "D"),
  ncol = 2, nrow = 2,
  label.x = 0.01, 
  label.y = 0.08 
)

compare_rich_shannon
#ggsave("output/compare_rich_shannon.png", width = 10, height = 10)

# abundance, richness and shannon diversity
compare_abun_rich_shannon <- ggarrange(
  plot_abundance_by_site_cacao_arrange + rremove("xlab"), 
  plot_abundance_by_site_inactive_arrange + rremove("xlab") + rremove("ylab"),
  plot_richness_by_site_cacao_arrange + rremove("xlab"), 
  plot_richness_by_site_inactive_arrange + rremove("ylab"),
  plot_shannon_by_site_cacao_arrange,
  plot_shannon_by_site_inactive_arrange + rremove("ylab"),
  labels = c("A", "B", "C", "D", "E", "F"),
  ncol = 2, nrow = 3,
  label.x = 0.015, #.01
  label.y = 0.13  #.13
)
compare_abun_rich_shannon

ggsave("output/compare_abun_rich_shannon_draft_5.png", width = 10, height = 10)
```


#3. Contextualize cacao sites against reference sites


To do this, need to use estimates of asymptotic species richness 
- b/c these sites did not have even sample coverage
- Compare rarefied specues richness across different landuse types
- use vegan package

create separate matrix for each landuse type
```{r}
# shade
head(data.shade)
# convert the data frame into a list matrix for inext
SH <- as.matrix(apply(data.shade[,-1],2, as.integer))
# site name as row names
row.names(SH) <- data.shade[,1]

# sun
head(data.sun)
SU <- as.matrix(apply(data.sun[,-1],2, as.integer))
row.names(SU) <- data.sun[,1]

# abandoned
head(data.aba)
AB <- as.matrix(apply(data.aba[,-1],2, as.integer))
row.names(AB) <- data.aba[,1]

# forest
head(data.for)
# # ### what happens if remove FLL-B dud site?
# data.for <- data.for[1:4,] 
FO <- as.matrix(apply(data.for[,-1],2, as.integer))
row.names(FO) <- data.for[,1]
```


##a. Species accumulation curve
```{r}
# Define assemblages as separate matrices
assemblages <- list(SHADE = SH, SUN = SU, ABANDON = AB, FOREST = FO)
colors <- c("#A6611A", "#DFC27D", "#80CDC1", "#018571")
labels <- c("Shade cacao", "Sun cacao", "Abandoned plantation", "Forest")

# Calculate species accumulation curves for each assemblage
accum_curves <- lapply(assemblages, 
                       function(x) specaccum(x, 
                                             method = "random", 
                                             permutations = 100))

# Determine x and y axis limits based on the maximum values from all accumulation curves
x_max <- max(sapply(accum_curves, function(curve) max(curve$sites)))
y_max <- max(sapply(accum_curves, function(curve) max(curve$richness)))

#### Create species accumulation plot with confidence intervals ####
# Create an empty plot with refined x and y limits for all assemblages
plot(NULL, xlim = c(0, x_max), ylim = c(0, y_max + 5),  # Add buffer to y-axis
     xlab = "Sample Size", ylab = "Species Richness",
     main = "Standardized species accumulation curves for each landuse type")

# Loop through each assemblage, add accumulation curve and CI to the plot
for (i in seq_along(accum_curves)) {
  # Extract the current curve
  curve <- accum_curves[[i]]
  
  # Plot the confidence interval as a shaded polygon
  polygon(c(curve$sites, rev(curve$sites)), 
          c(curve$richness + curve$sd, rev(curve$richness - curve$sd)),
          col = adjustcolor(colors[i], alpha.f = 0.2), border = NA)
  
  # Plot the accumulation curve with lines
  lines(curve$sites, curve$richness, col = colors[i], lwd = 2)
}

# Add a legend to the plot (outside the loop)
legend("bottomright", legend = labels, col = colors, lty = 1, lwd = 2, cex = 0.8)
```


##b. Rarefaction

Technique to assess expected species richness
- allows calculation of species richness for given number of samples based on
  construction of rarefaction curves
  
issue is that the larger number of ind sampled, more species will be found
- rarefaction curves created by randomly re-sampling pool of N samples multiple
  times and then plotting average number of species found in each sample
- rarefaction generates expected number of species in small collection of n ind
  drawn from large pool of N samples
- typically grow rapidly at first, then slowly as only rarest species remain to
  be sampled

use rarefy and rarecurve functions
```{r}
# define rarefaction depth - smallest total abundance from four matrices
min_sample_size <- min(rowSums(SH), rowSums(SU), rowSums(AB), rowSums(FO))

# perform the rarefaction - expected species richness if each sample standardized to min
rarefied_richness_shade <- rarefy(SH, min_sample_size)
rarefied_richness_sun <- rarefy(SU, min_sample_size)
rarefied_richness_abandon <- rarefy(AB, min_sample_size)
rarefied_richness_forest <- rarefy(FO, min_sample_size)

# combine results and calculate means for each land use type for easy comparison:
mean_rarefied_richness <- c(mean(rarefied_richness_shade), 
                   mean(rarefied_richness_sun), 
                   mean(rarefied_richness_abandon), 
                   mean(rarefied_richness_forest))

# visualize simple plot
boxplot(rarefied_richness_shade, 
        rarefied_richness_sun, 
        rarefied_richness_abandon, 
        rarefied_richness_forest,
        names = c("Shade cacao", "Sun cacao", "Abandoned cacao", "Forest"),
        ylab = "Rarefied Species Richness")
```

not totally necessary, but interesting to see dist. of ind at site and sp level
```{r}
# abundances for each site
site_abun <- rowSums(new_mat)
site_abun

# abundances for each species
sp_abun <- colSums(new_mat)
sp_abun
```

run rarefaction with all sites
- this is the default
```{r}
raremin <- min(rowSums(new_mat))
raremin # min ind in a site is 6 from VC-C

# run rarefaction with all sites (including super low site VC-C)
sRare <- rarefy(new_mat, raremin)
sRare

# join diversity_by_site data with sRare in a df
diversity_by_site <- sRare %>% 
  enframe() %>% 
  full_join(diversity_by_site, by = c("name" = "Sitio")) %>%
  rename(sRare = value)

# Anova for rarefied species richness by tpe
sRare_aov <- aov(sRare ~ Tipo, data = diversity_by_site)
summary(sRare_aov)
# p-value is 0.266, not sig difference
```


rarefaction uses smallest number of obs per sample to extrapolate expected
number if all other samples only had that number of obs
- VC-C only had 6!!! This certainly impacts the results of rarefaction
- try running rarefaction analysis without this site, see how this impacts things
- of course present this caveat!
```{r}
# remove VC-C from new_mat matrix via row position
new_mat_abundant <- new_mat %>%
  filter(!row_number() %in% c(23))
new_mat_abundant # confirmed VC-C is absent

# let's see what happens with this as our raremin - now set at 22
raremin_abundant <- min(rowSums(new_mat_abundant))
raremin_abundant

# expected rarefied species richness if only 22 ind collected at each site
sRare_abundant <- rarefy(new_mat_abundant, raremin_abundant)
sRare_abundant

# remove VC-C from diversity_by_site to ensure full_join happens
diversity_by_site_abundant <- diversity_by_site %>%
  filter(!name == 'VC-C')

# join diversity_by_site_abundant with sRare_abundant in a df
diversity_by_site_abundant <- sRare_abundant %>% 
  enframe() %>% 
  left_join(diversity_by_site_abundant, by = "name") %>%
  rename(sRare_abundant = value)

# Anova for rarefied species richness by type without VC-C
sRare_abundant_aov <- aov(sRare_abundant ~ Tipo, 
                          data = diversity_by_site_abundant)
summary(sRare_abundant_aov) # p-value 0.205, not sig difference
```

no difference in findings based on inclusion / exclusion of VC-C, so include it
```{r}
# site labels for 4 site comparison
site_labels <-  c("Shade Cacao \n (n = 12)", 
                  "Sun Cacao \n (n = 11)", 
                  "Abandoned Plantation \n (n = 4)",
                  "Remnant Forest \n (n = 5)")

# reorder factor levels to match site_labels
diversity_by_site$Tipo <- factor(diversity_by_site$Tipo,
                         c("N", "C", "V", "B"))

# plot rarefied species richness across landuse types using palet, my_theme
plot_rare_by_type <- ggplot(diversity_by_site, 
                                     aes(x = Tipo, y = sRare, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Land use type",
       y = "Rarefied species richness")

plot_rare_by_type
#ggsave("output/plot_rare_by_type.png", width = 10, height = 10)
```


##c. Extrapolation

use iNEXT to make rarefaction curves of each site type
- this figure doesn't have extrapolation - vegan is doing the same thing above
```{r}
#load matrix
attach(MATFROGH)
data <- MATFROGH

# Check the structure of your matrix
str(MATFROGH)

# Ensure Type is a factor
MATFROGH$Type <- as.factor(MATFROGH$Type)

# Summarize abundances for each Type and split into a list
abundance_list <- MATFROGH %>%
  group_by(Type) %>%
  summarise(across(where(is.numeric), sum)) %>%  # Summarize numeric columns
  split(.$Type)  # Split the summarized data by Type

# Remove the Type column within each list element
abundance_list <- lapply(abundance_list, function(x) x[, -1])  # Drop 'Type' column

# Summarize abundances for each Type and split into numeric vectors
abundance_list <- MATFROGH %>%
  group_by(Type) %>%
  summarise(across(where(is.numeric), sum)) %>%
  split(.$Type)  # Split by Type

# Convert each element of the list to a numeric vector
abundance_list <- lapply(abundance_list, 
                         function(x) as.numeric(x[, -1]))  # Drop the 'Type' column

# Check the structure of abundance_list
str(abundance_list)

###RAREFACTION###
rare_results <- iNEXT(abundance_list, q = 0, datatype = "abundance")

rare_results
# Plot the rarefaction curves

ggiNEXT(rare_results, type = 1) +
  ggtitle("Rarefaction Curves by Habitat Type") +
  xlab("Sample Size") +
  ylab("Species Richness")
```

## WORKING

use iNEXT to extrapolate rarefaction curve to see what expected values would be
- visually compare extrapolated rarefied species richness across landuse types
- calculate extrapolated rarefied species richness
```{r}
###VISUALIZATION###
# Perform rarefaction analysis with extended extrapolation
rare_results_extended <- iNEXT(abundance_list, q = 0, 
                               datatype = "abundance", endpoint = 2000)

# Plot rarefaction/extrapolation curves with extended extrapolation
ggiNEXT(rare_results_extended, type = 1) +
  ggtitle("Extended Rarefaction and Extrapolation Curves by Habitat Type") +
  xlab("Sample Size") +
  ylab("Species Richness")



extrapolated_rich_plot <- ggiNEXT(rare_results_extended, type = 1) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Sample size",
       y = "Rarefied species richness")

extrapolated_rich_plot

###############################################################################
### THIS IS SO SO SO CLOSE! JUST NEED TO CHANGE DISPLAY OF VALUES!!!!!

# working:
ggiNEXT(rare_results_extended, type = 1) +
  ggtitle("Extended Rarefaction and Extrapolation Curves by Habitat Type") +
  xlab("Sample Size") +
  ylab("Species Richness") +
  theme_minimal() +
  scale_color_manual(values = c("ABANDON" = "#80CDC1", 
                                "FOREST" = "#018571", 
                                "SHADE" = "#A6611A",
                                "SUN" = "#DFC27D")) 

###############################################################################

# Extract the iNextEst data frame
extrapolated_richness <- rare_results_extended$iNextEst

# Check the structure
str(extrapolated_richness)

# View the data
head(extrapolated_richness)

# Extract size-based extrapolated richness values
extrapolated_summary <- extrapolated_richness$size_based %>%
  filter(Method == "Extrapolation", m == 2000) %>%
  select(Assemblage, m, qD, qD.LCL, qD.UCL)

# Print the summary
print(extrapolated_summary)
```








##d. Endangered species
```{r}
threatened_abundance_by_site <- new_threatened %>%
  group_by(Tipo, Final_ID) %>%
  summarize(num_ind = n_distinct(Numero))

threatened_abundance_by_site
```


##e. Dissimilarity indices
```{r}
# Create a site type vector
site_types <- c(rep("Shade cacao", 12), 
                rep("Sun cacao", 11), 
                rep("Abandoned plantation", 4), 
                rep("Remnant forest", 5))

dissimilarity_matrix <- vegdist(new_mat, method = "bray")

# NMDS - used by Mendenhall!
nmds <- metaMDS(dissimilarity_matrix, k = 2) # stress of .08
plot(nmds)

# PCA - (if dissimilarities are not too high-dimensional)
pca <- cmdscale(dissimilarity_matrix, k = 2)
plot(pca[, 1], pca[, 2], xlab = "PC1", ylab = "PC2")

# visualize and color by site type
nmds_data <- as.data.frame(scores(nmds))
nmds_data$site_type <- factor(site_types, 
                              levels = c("Shade cacao", "Sun cacao", 
                                         "Abandoned plantation", "Remnant forest"))

# plot with ggplot, color by site type
plot_nmds_bray <- ggplot(nmds_data, aes(x = NMDS1, y = NMDS2, 
                                        color = site_type)) +
  geom_point(size = 4) +
  scale_color_manual(values = c("Shade cacao" = "#A6611A", 
                               "Sun cacao" = "#DFC27D", 
                               "Abandoned plantation" = "#80CDC1", 
                               "Remnant forest" = "#018571")) +
  theme_minimal() +
  labs(color = "Land use type")

plot_nmds_bray
#ggsave("output/plot_nmds_bray.png", width = 10, height = 10)
```


#4. Multiple regression for drivers of amphibian abundance within plantations

Collected a series of local (quad-level), landscape (site-level) variables
- how do these explain the abundance of amphibians at a transect-level?

Start at transect level
- much greater resolution for local variables
- greater statistical power (4 x 23 transects)
- abundance have high enough numbers to work at transect level (not tru for sp)

##a. PCA
- can only work for continuous variables

standard PCA fitting and plotting
```{r}
pca_fit <- prcomp(
  x = trans_richness_local[, -c(1,2,5,6,7,9,10,11,12,13,14,15)], # remove these columns
                  scale. = TRUE)

pca_summary <- summary(pca_fit)
pca_summary # first two axes explain 55% of variance

ls(pca_summary) # Lists the objects produced by summary

pca_summary$importance

# biplot - standard, kinda ugly!
biplot(x = pca_fit)
```

fancy biplot with specified colors - need to map manually
```{r}
# Define specific colors
colors <- c("#DFC27D", "#A6611A")
names(colors) <- levels(as.factor(site_richness_local$Tipo))

# Mapping site type abbreviations to full names
site_type_names <- c(C = "Sun cacao", N = "Shade cacao")

# Map colors to points based on Tipo
point_colors <- colors[as.factor(site_richness_local$Tipo)]

# Extract PCA components and loadings
pca_scores <- pca_fit$x  # Observation scores
pca_loadings <- pca_fit$rotation  # Variable loadings

# Create the biplot
plot(pca_scores[, 1], pca_scores[, 2], 
     col = point_colors, 
     pch = 19, 
     xlab = "PC1", 
     ylab = "PC2", 
     main = "Customized PCA Biplot",
     xlim = c(min(c(pca_scores[, 1], pca_loadings[, 1])) * 1.2, 
              max(c(pca_scores[, 1], pca_loadings[, 1])) * 1.2),
     ylim = c(min(c(pca_scores[, 2], pca_loadings[, 2])) * 1.2, 
              max(c(pca_scores[, 2], pca_loadings[, 2])) * 1.2))

# Add arrows for variable loadings
arrows(0, 0, pca_loadings[, 1] * max(abs(pca_scores[, 1])), 
       pca_loadings[, 2] * max(abs(pca_scores[, 2])), 
       length = 0.1, col = "black")

# Add variable labels
text(pca_loadings[, 1] * max(abs(pca_scores[, 1])), 
     pca_loadings[, 2] * max(abs(pca_scores[, 2])), 
     labels = rownames(pca_loadings), 
     col = "black", cex = 0.8)

# Add legend
legend("bottomright", 
       legend = c("Shade", "Sun"), 
       col = c("#A6611A", "#DFC27D"), 
       pch = 19)
```

use ggplot2 for publication level biplot
```{r}
# Prepare data for ggplot
data_pca <- data.frame(PC1 = pca_scores[, 1],
                       PC2 = pca_scores[, 2],
                       Group = as.factor(trans_richness_local$Tipo))

loading_data <- data.frame(PC1 = pca_loadings[, 1] * max(abs(pca_scores[, 1])),
                           PC2 = pca_loadings[, 2] * max(abs(pca_scores[, 2])),
                           Variable = rownames(pca_loadings))

# Custom labels for legend
group_labels <- c("C" = "Sun", "N" = "Shade")

# Create plot
tran_pca_plot <- ggplot() +
  # Plot points with group coloring
  geom_point(data = data_pca, aes(x = PC1, y = PC2, color = Group), size = 3) +
  # Add arrows for variable loadings
  geom_segment(data = loading_data, aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  # Add labels to the arrows
  geom_text_repel(data = loading_data, aes(x = PC1, y = PC2, label = Variable),
                  color = "black", size = 4) +
  # Customize color scale and legend
  scale_color_manual(values = c("#DFC27D", "#A6611A"),
                     labels = group_labels) +
  labs(x = "PC1", y = "PC2", color = "Cacao type") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
  )

tran_pca_plot
#ggsave("output/tran_pca_plot.png", width = 10, height = 10)
```

Plot first two axes of PCA, no variable loadings
```{r}
# Define specific colors
colors <- c("#DFC27D", "#A6611A")
names(colors) <- levels(as.factor(trans_richness_local$Tipo))

# Mapping site type abbreviations to full names
site_type_names <- c(C = "Sun cacao", N = "Shade cacao")

# Scatterplot with specified colors
plot(x = pca_fit$x[, 1],
     y = pca_fit$x[, 2],
     xlab = "PC 1 (Explained variance: 30%)",
     ylab = "PC 2 (Explained variance: 24%)",
     col = colors[as.factor(site_richness_local$Tipo)],
     pch = 19,
     main = "PCA Scatterplot")

# Add legend with full names
legend("topleft", 
       legend = site_type_names[levels(as.factor(site_richness_local$Tipo))], 
       col = colors, 
       pch = 19)
```

```{r}
pca_fit <- prcomp(
  x = trans_richness_local[, -c(1,2,5,6,7,9,10,11,12,13,14,15)], # remove these columns
                  scale. = TRUE)

pca_summary <- summary(pca_fit)
pca_summary # first two axes explain 52% of variation

ls(pca_summary) # Lists the objects produced by summary

pca_summary$importance

# biplot - kinda ugly!
biplot(x = pca_fit)

# just plot first two axes of PCA
plot(x = pca_fit$x[, 1],
     y = pca_fit$x[, 2],
     xlab = "PC 1",
     ylab = "PC 2")
```


##b. Correlation

need to address potential variable correlation before building regression models

visually assess potential correlation of scaled variables in trans_richness_local
- tran_mean_hoja_scaled and tran_mean_CC_scaled are highly correlated!
- this makes sense as more leaf coverage above should translate to more below
- so only use one of these variables in model creation
```{r}
# would only expect local vars or landscape vars to be correlated with each other
ggscatter(trans_richness_local, x = "tran_mean_hoja_scaled", 
          y = "tran_mean_veg_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "tran_mean_hoja_scaled", ylab = "tran_mean_veg_scaled")
# r = -.21

ggscatter(trans_richness_local, x = "tran_mean_hoja_scaled", 
          y = "tran_mean_CC_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "tran_mean_hoja_scaled", ylab = "tran_mean_CC_scaled")
# r = .7 HIGHLY SCALED

ggscatter(trans_richness_local, x = "tran_mean_veg_scaled", 
          y = "tran_mean_CC_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "tran_mean_veg_scaled", ylab = "tran_mean_CC_scaled")
# r = .0001

ggscatter(trans_richness_local, x = "water_dist_scaled", 
          y = "forest_cover_400_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "water_dist_scaled", ylab = "forest_cover_400_scaled")
# r = -.42
```

confirm this with correlation matrix
- tran_mean_CC_scaled and tran_mean_hoja_scaled highly correlated (.7)
```{r}
cor(trans_richness_local[, c("tran_mean_hoja_scaled", 
                             "tran_mean_CC_scaled", 
                             "water_dist_scaled", 
                             "forest_cover_400_scaled",
                             "tran_mean_veg_scaled")])
```

create pca variable based off CC and hoja b/c highly correlated
- probably just end up using tran_mean_hoja but still keep just in case
```{r}
pca_CC_hoja <- prcomp(trans_richness_local
              [, c("tran_mean_hoja_scaled", "tran_mean_CC_scaled")], scale. = TRUE)
trans_richness_local$pca_CC_hoja <- pca_CC_hoja$x[,1]
```

##c. Drivers of transect-level amphibian abundance
use glmer package to build full multiple regression
- allows for inclusion of both random and rixed effects 
- use canopy cover instead of hoja, no quantile deviation detected
- work at transect level

full model
```{r}
############# FULL MODEL FITS - AIC 618.5 - CC sig pos ##############
# full model with glmer, CC, veg as R.E., poisson
modfull_abundance <- glmer(num_ind ~ tran_mean_CC_scaled + # use CC not hoja
                   water_dist_scaled +
                   forest_cover_400_scaled +
                   (1 | tran_mean_veg_scaled) +
                   (1 | Sitio),
              data = trans_richness_local,
              family = poisson())

summary(modfull_abundance) # tran_mean_CC_sig - AIC 618.5
vif(modfull_abundance) # vif not a concern
check_overdispersion(modfull_abundance) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = modfull_abundance)
plot(simulationOutput)

testDispersion(simulationOutput) # not over or under dispersed
testZeroInflation(simulationOutput) # zero inflation not required
############# FULL MODEL FITS - AIC 618.5 - CC sig pos ##############

# quantile deviation detected when using hoja instead of CC
# model fit errors when using veg as a fixed effect instead of a random effect
# quantile deviation detected when using negative binomial instead of poisson
```

leave one out cross-validation for candidate models for abundance at transect-level
- start from full model which confirmed fit model assumptions
```{r}
# STEP 1 - remove one predictor variable


############# MODEL NO CC COMBINED ADJUSTED QUANTILE TEST SIG ##############
# full model with glmer, no CC, veg as R.E., poisson
mod_abundance_no_CC <- glmer(num_ind ~ water_dist_scaled +
                   forest_cover_400_scaled +
                   (1 | tran_mean_veg_scaled) +
                   (1 | Sitio),
              data = trans_richness_local,
              family = poisson())

summary(mod_abundance_no_CC) # nothing sig, AIC 629.1
vif(mod_abundance_no_CC) # vif not a concern
check_overdispersion(mod_abundance_no_CC) # not overdispersed

# assess model fit using dHARMA
# CRAZY! deviation sig, combined adjusted quantile test sig
simulationOutput <- simulateResiduals(fittedModel = mod_abundance_no_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not over or under dispersed
testZeroInflation(simulationOutput) # zero inflation not required


############# MODEL NO WATER FITS - AIC 616.5 - CC sig pos ##############
# full model with glmer, no water, veg as R.E., poisson
mod_abundance_no_water <- glmer(num_ind ~ tran_mean_CC_scaled +
                   forest_cover_400_scaled +
                   (1 | tran_mean_veg_scaled) +
                   (1 | Sitio),
              data = trans_richness_local,
              family = poisson())

summary(mod_abundance_no_water) # CC sig, AIC 616.5
vif(mod_abundance_no_water) # vif not a concern
check_overdispersion(mod_abundance_no_water) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_abundance_no_water)
plot(simulationOutput)

testDispersion(simulationOutput) # not over or under dispersed
testZeroInflation(simulationOutput) # zero inflation not required


###### MODEL NO FOREST - Quant dev detected (adjusted quantile test n.s) #######
#- AIC 616.5 - CC sig pos #
# full model with glmer, no forest, veg as R.E., poisson
mod_abundance_no_forest <- glmer(num_ind ~ tran_mean_CC_scaled +
                   water_dist_scaled +
                   (1 | tran_mean_veg_scaled) +
                   (1 | Sitio),
              data = trans_richness_local,
              family = poisson())

summary(mod_abundance_no_forest) # CC sig, AIC 616.5
vif(mod_abundance_no_forest) # vif not a concern
check_overdispersion(mod_abundance_no_forest) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_abundance_no_forest)
plot(simulationOutput)

testDispersion(simulationOutput) # not over or under dispersed
testZeroInflation(simulationOutput) # zero inflation not required


# STEP 2 - remove 2nd predictor variable


####### MODEL NO WATER NO FOREST - Quant dev detected (adjusted quantile test sig) ######
#AIC 614.5 - CC sig pos
# full model with glmer, no water, no forest, veg as R.E., poisson
mod_abundance_no_water_forest <- glmer(num_ind ~ tran_mean_CC_scaled +
                   (1 | tran_mean_veg_scaled) +
                   (1 | Sitio),
              data = trans_richness_local,
              family = poisson())

summary(mod_abundance_no_water_forest) # CC sig, AIC 614.5
# vif(mod_abundance_no_water_forest) # ERROR: model contains fewer than two terms
check_overdispersion(mod_abundance_no_water_forest) # not overdispersed

# assess model fit using dHARMA - 
simulationOutput <- simulateResiduals(fittedModel = mod_abundance_no_water_forest)
plot(simulationOutput)

testDispersion(simulationOutput) # not over or under dispersed
testZeroInflation(simulationOutput) # zero inflation not required


####### MODEL NO WATER NO CC - Quant dev detected (adj quant test sig) ######
# AIC 627.5 - no sig
# full model with glmer, no water, no CC, veg as R.E., poisson
mod_abundance_no_water_CC <- glmer(num_ind ~ forest_cover_400_scaled +
                   (1 | tran_mean_veg_scaled) +
                   (1 | Sitio),
              data = trans_richness_local,
              family = poisson())

summary(mod_abundance_no_water_CC) # no sig, AIC 627.5
# vif(mod_abundance_no_water_CC) # ERROR: model contains fewer than two terms
check_overdispersion(mod_abundance_no_water_CC) # not overdispersed

# assess model fit using dHARMA 
# CRAZY! Quant dev detected, combined adjusted quant test sig
simulationOutput <- simulateResiduals(fittedModel = mod_abundance_no_water_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not over or under dispersed
testZeroInflation(simulationOutput) # zero inflation not required


# STOP HERE - none of Step 2 models fit
```

the best fitting candidate model for amphibian abundance contains:
- local canopy cover as a significant positive fixed effect
- surrounding landscape forest cover as a fixed effect
- Site and local herbaceous vegetation cover as random effects


#5. Multiple regression for drivers of amphibian richness in plantations

Drivers of site-level species richness
use glmer package to build multiple regression models
- allows for inclusion of both random and fixed effects 

Work at site-level to improve model fit
- At transect-level many sites have 1-2 species
- Less statistical power at site-level (23 sites), but better understanding of diversity
- Start with the predictor models that were important for transect-level abundance

##a. PCA
- can only work for continuous variables

standard PCA fitting and plotting
```{r}
pca_fit <- prcomp(
  x = site_richness_local[, -c(1,4,5,6,7,10,11,12,16)], # remove these columns
                  scale. = TRUE)

pca_summary <- summary(pca_fit)
pca_summary # first two axes explain 53% of variance

ls(pca_summary) # Lists the objects produced by summary

pca_summary$importance

# biplot - standard, kinda ugly!
biplot(x = pca_fit)
```

fancy biplot with specified colors - need to map manually
```{r}
# Define specific colors
colors <- c("#DFC27D", "#A6611A")
names(colors) <- levels(as.factor(site_richness_local$Tipo))

# Mapping site type abbreviations to full names
site_type_names <- c(C = "Sun cacao", N = "Shade cacao")

# Map colors to points based on Tipo
point_colors <- colors[as.factor(site_richness_local$Tipo)]

# Extract PCA components and loadings
pca_scores <- pca_fit$x  # Observation scores
pca_loadings <- pca_fit$rotation  # Variable loadings

# Create the biplot
plot(pca_scores[, 1], pca_scores[, 2], 
     col = point_colors, 
     pch = 19, 
     xlab = "PC1", 
     ylab = "PC2", 
     main = "Customized PCA Biplot",
     xlim = c(min(c(pca_scores[, 1], pca_loadings[, 1])) * 1.2, 
              max(c(pca_scores[, 1], pca_loadings[, 1])) * 1.2),
     ylim = c(min(c(pca_scores[, 2], pca_loadings[, 2])) * 1.2, 
              max(c(pca_scores[, 2], pca_loadings[, 2])) * 1.2))

# Add arrows for variable loadings
arrows(0, 0, pca_loadings[, 1] * max(abs(pca_scores[, 1])), 
       pca_loadings[, 2] * max(abs(pca_scores[, 2])), 
       length = 0.1, col = "black")

# Add variable labels
text(pca_loadings[, 1] * max(abs(pca_scores[, 1])), 
     pca_loadings[, 2] * max(abs(pca_scores[, 2])), 
     labels = rownames(pca_loadings), 
     col = "black", cex = 0.8)

# Add legend
legend("bottomright", 
       legend = c("Shade", "Sun"), 
       col = c("#A6611A", "#DFC27D"), 
       pch = 19)
```

use ggplot2 for publication level biplot
```{r}
# Prepare data for ggplot
data_pca <- data.frame(PC1 = pca_scores[, 1],
                       PC2 = pca_scores[, 2],
                       Group = as.factor(site_richness_local$Tipo))

loading_data <- data.frame(PC1 = pca_loadings[, 1] * max(abs(pca_scores[, 1])),
                           PC2 = pca_loadings[, 2] * max(abs(pca_scores[, 2])),
                           Variable = rownames(pca_loadings))

# Custom labels for legend
group_labels <- c("C" = "Sun", "N" = "Shade")

# Create plot
site_pca_plot <- ggplot() +
  # Plot points with group coloring
  geom_point(data = data_pca, aes(x = PC1, y = PC2, color = Group), size = 3) +
  # Add arrows for variable loadings
  geom_segment(data = loading_data, aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  # Add labels to the arrows
  geom_text_repel(data = loading_data, aes(x = PC1, y = PC2, label = Variable),
                  color = "black", size = 4) +
  # Customize color scale and legend
  scale_color_manual(values = c("#DFC27D", "#A6611A"),
                     labels = group_labels) +
  labs(x = "PC1", y = "PC2", color = "Cacao type") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
  )

site_pca_plot
#ggsave("output/site_pca_plot.png", width = 10, height = 10)
```

Plot first two axes of PCA, no variable loadings
```{r}
# Define specific colors
colors <- c("#DFC27D", "#A6611A")
names(colors) <- levels(as.factor(site_richness_local$Tipo))

# Mapping site type abbreviations to full names
site_type_names <- c(C = "Sun cacao", N = "Shade cacao")

# Scatterplot with specified colors
plot(x = pca_fit$x[, 1],
     y = pca_fit$x[, 2],
     xlab = "PC 1 (Explained variance: 30%)",
     ylab = "PC 2 (Explained variance: 24%)",
     col = colors[as.factor(site_richness_local$Tipo)],
     pch = 19,
     main = "PCA Scatterplot")

# Add legend with full names
legend("bottomright", 
       legend = site_type_names[levels(as.factor(site_richness_local$Tipo))], 
       col = colors, 
       pch = 19)
```

##b. Correlation

need to address potential variable correlation before building regression models
- same as transect level! site_mean_hoja and site_mean_CC are highly correlated
- so only use one when building models instead of both

visually assess potential correlation of scaled variables in site_richness_local
```{r}
# would only expect local vars or landscape vars to be correlated with each other
ggscatter(site_richness_local, x = "site_mean_hoja_scaled", 
          y = "site_mean_veg_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "site_mean_hoja_scaled", ylab = "site_mean_veg_scaled")
# r = -.17

ggscatter(site_richness_local, x = "site_mean_hoja_scaled", 
          y = "site_mean_CC_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "site_mean_hoja_scaled", ylab = "site_mean_CC_scaled")
# r = .79 HIGHLY SCALED

ggscatter(site_richness_local, x = "site_mean_veg_scaled", 
          y = "site_mean_CC_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "site_mean_veg_scaled", ylab = "site_mean_CC_scaled")
# r = .086

ggscatter(site_richness_local, x = "water_dist_scaled", 
          y = "forest_cover_400_scaled", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "water_dist_scaled", ylab = "forest_cover_400_scaled")
# r = -.41
```

confirm this with correlation matrix
- site_mean_CC_scaled and site_mean_hoja_scaled highly correlated (.79)
```{r}
cor(site_richness_local[, c("site_mean_hoja_scaled", 
                             "site_mean_CC_scaled", 
                             "water_dist_scaled", 
                             "forest_cover_400_scaled",
                             "site_mean_veg_scaled")])
```

create pca variable based off CC and hoja b/c highly correlated
- probably just end up using site_mean_hoja_scaled but still keep just in case
```{r}
pca_CC_hoja <- prcomp(site_richness_local
              [, c("site_mean_hoja_scaled", "site_mean_CC_scaled")], scale. = TRUE)
site_richness_local$pca_CC_hoja <- pca_CC_hoja$x[,1]
```

##c. Drivers of site-level amphibian richness
use glm package to build full multiple regression
- inclusion of random effects produced singular fit errors with glmer
- use canopy cover instead of hoja b/c highly correlated
- model needs to be simplified b/c less power and replication at site-level
- work at site level b/c trans level richness counts are not as meaningful

full model
```{r}
############# FULL MODEL - nothing sig - AIC 103.5 ##############
# full model with glm, no R.E., poisson
mod_richness_full <- glm(no_species ~ site_mean_CC_scaled +
                                    site_mean_veg_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_full) # nothing sig - AIC 103.5
vif(mod_richness_full) # VIF not concern
check_overdispersion(mod_richness_full) # not over dispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated
############# FULL MODEL - nothing sig - AIC 103.5 ##############
```

leave one out cross-validation for candidate models for richness at site-level


Step 1:
- start from full model which confirmed fit model assumptions
- full model AIC was 103.5
```{r}
############# STEP 1 - remove one predictor variable ################


############# FULL MODEL NO CC - AIC 101.51 - nothing sig ##############
# full model with glm, no CC, no R.E., poisson
mod_richness_full_no_CC <- glm(no_species ~ 
                                    # site_mean_CC_scaled +
                                    site_mean_veg_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_full_no_CC) # nothing sig - AIC 101.51
vif(mod_richness_full_no_CC) # vif not a concern
check_overdispersion(mod_richness_full_no_CC) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full_no_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

############# FULL MODEL NO VEG - AIC 101.48 - nothing sig ##############
# full model with glm, no veg, no R.E., poisson
mod_richness_full_no_veg <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    # site_mean_veg_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_full_no_veg) # nothing sig - AIC 101.48
vif(mod_richness_full_no_veg) # vif not a concern
check_overdispersion(mod_richness_full_no_veg) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full_no_veg)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

####### FULL MODEL NO WATER - QUANT DEVIATIONS DETECTED (test n.s) ######
# nothing sig - AIC 101.52
# full model with glm, no water, no R.E., poisson
mod_richness_full_no_water <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    site_mean_veg_scaled +
                                    # water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_full_no_water) # nothing sig - AIC 101.52
vif(mod_richness_full_no_water) # vif not a concern
check_overdispersion(mod_richness_full_no_water) # not overdispersed

# assess model fit using dHARMA - QUANTILE DEVIATIONS DETECTED
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full_no_water)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

############# FULL MODEL NO FOREST - nothing sig - AIC 101.91 ##############
# full model with glm, no forest, no R.E., poisson
mod_richness_full_no_forest <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    site_mean_veg_scaled +
                                    water_dist_scaled +
                                    # forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())


summary(mod_richness_full_no_forest) # nothing sig - AIC 101.91
vif(mod_richness_full_no_forest) # vif not a concern
check_overdispersion(mod_richness_full_no_forest) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full_no_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

###### FULL MODEL NO ELEVATION - QUANT DEVIATIONS DETECTED (test n.s.) ######
# nothing sig - AIC 102.96
# full model with glm, no forest, no R.E., poisson
mod_richness_full_no_elevation <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    site_mean_veg_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled, # +
                                    # Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_full_no_elevation) # nothing sig - AIC 102.96
vif(mod_richness_full_no_elevation) # vif not a concern
check_overdispersion(mod_richness_full_no_elevation) # not overdispersed

# assess model fit using dHARMA - QUANTILE DEVIATIONS DETECTED (quantile test n.s.)
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full_no_elevation)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated
```


Step 2:
- start from no veg model which confirmed fit model assumptions
- no veg model AIC was 101.48
```{r}
############# STEP 2 - remove second predictor variable ############
# start with NO VEG model as this had lowest AIC from Step 1


############# MODEL NO VEG, CC - outer Newton didn't converge? ##############
# nothing sig - AIC 99.51
# model without veg, CC, glm, no R.E., poisson
mod_richness_no_veg_CC <- glm(no_species ~ 
                                    # site_mean_CC_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_CC) # nothing sig - AIC 99.51
vif(mod_richness_no_veg_CC) # vif not a concern
check_overdispersion(mod_richness_no_veg_CC) # not overdispersed

# assess model fit using dHARMA - outer Newton did not converge fully?
simulationOutput <- simulateResiduals(fittedModel = mod_richness_no_veg_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

############# MODEL NO VEG, WATER - nothing sig - AIC 99.52 ##############
# model without veg, water, glm, no R.E., poisson
mod_richness_no_veg_water <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    # water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_water) # nothing sig - AIC 99.52
vif(mod_richness_no_veg_water) # vif not a concern
check_overdispersion(mod_richness_no_veg_water) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_no_veg_water)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

############# MODEL NO VEG, FOREST - nothing sig - AIC 99.975 ##############
# model without veg, forest, glm, no R.E., poisson
mod_richness_no_veg_forest <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    water_dist_scaled +
                                    # forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_forest) # nothing sig - AIC 99.975
vif(mod_richness_no_veg_forest) # vif not a concern
check_overdispersion(mod_richness_no_veg_forest) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_no_veg_forest)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated

############# MODEL NO VEG, ELEVATION - nothing sig - AIC 101.04 ##############
# model without veg, elevation, glm, no R.E., poisson
mod_richness_no_veg_elevation <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled, #+
                                    # Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_elevation) # nothing sig - AIC 101.04
vif(mod_richness_no_veg_elevation) # vif not a concern
check_overdispersion(mod_richness_no_veg_elevation) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_no_veg_elevation)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated
```


Step 3:
- start from no veg no water model which confirmed fit model assumptions
- no veg no water model AIC was 99.52
```{r}
############# STEP 3 - remove third predictor variable ############
# start with NO VEG NO WATER model as this had lowest AIC from Step 2


###### MODEL NO VEG, WATER, CC - QUANT DEVIATIONS DETECTED (test n.s.) #######
# nothing sig - AIC 97.53
# model without veg, water, CC, glm, no R.E., poisson
mod_richness_no_veg_water_CC <- glm(no_species ~ 
                                    # site_mean_CC_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_water_CC) # nothing sig AIC 97.53
vif(mod_richness_no_veg_water_CC) # vif not a concern
check_overdispersion(mod_richness_no_veg_water_CC) # not overdispersed

# assess model fit using dHARMA - QUANTILE DEVIATIONS DETECTED (QUANT TEST NS)
simulationOutput <- simulateResiduals(fittedModel = mod_richness_no_veg_water_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not overdispersed
testZeroInflation(simulationOutput) # zero inflated not necessary


###### MODEL NO VEG, WATER, FOREST - nothing sig - AIC 97.97 ########
# model without veg, water, forest glm, no R.E., poisson
mod_richness_no_veg_water_forest <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    # forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_water_forest) # nothing sig - AIC 97.975
vif(mod_richness_no_veg_water_forest) # vif not a concern
check_overdispersion(mod_richness_no_veg_water_forest) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_no_veg_water_forest)
plot(simulationOutput)

testDispersion(simulationOutput) #  not overdispersed
testZeroInflation(simulationOutput) # zero inflation not necessary


####### MODEL NO VEG, WATER, Elevation - nothing sig - AIC 99.07 ########
# model without veg, water, elevation glm, no R.E., poisson
mod_richness_no_veg_water_elevation <- glm(no_species ~ 
                                    site_mean_CC_scaled +
                                    forest_cover_400_scaled, #+
                                    # Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_no_veg_water_elevation) # nothing sig - AIC 99.07
vif(mod_richness_no_veg_water_elevation) # vif not a concern
check_overdispersion(mod_richness_no_veg_water_elevation) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(
  fittedModel = mod_richness_no_veg_water_elevation)
plot(simulationOutput)

testDispersion(simulationOutput) # not overdispersed
testZeroInflation(simulationOutput) # zero inflation not needed
```


Step 4:
- start from no veg no water no forest model which confirmed fit model assumptions
- no veg no water no forest model AIC was 97.97
- so now just looking at either elevation or CC
```{r}
############# STEP 4 - remove fourth predictor variable ############
# start with NO VEG NO WATER NO FOREST model as this had lowest AIC from Step 3
# only two predictor variables left :) easier to just name the var remaining


###### MODEL W/ ONLY ELEVATION - nothing sig - AIC 96.06 ########
# model with only elevation, glm, no R.E., poisson
mod_richness_only_elevation <- glm(no_species ~ 
                                    # site_mean_CC_scaled +
                                    Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_only_elevation) # nothing sig - AIC 96.06
# vif(mod_richness_only_elevation) # ERROR: model contains fewer than 2 terms
check_overdispersion(mod_richness_only_elevation) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_only_elevation)
plot(simulationOutput)

testDispersion(simulationOutput) # not overdispersed
testZeroInflation(simulationOutput) # not zero inflated


###### MODEL W/ ONLY CC - nothing sig - AIC 97.1 ########
# model with only cc, glm, no R.E., poisson
mod_richness_only_CC <- glm(no_species ~ 
                                    site_mean_CC_scaled, #+
                                    # Elevation,
                                    data = site_richness_local,
                                    family = poisson())

summary(mod_richness_only_CC) # nothing sig - AIC 97.1
# vif(mod_richness_only_CC) # ERROR: model contains fewer than 2 terms
check_overdispersion(mod_richness_only_CC) # not overdispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_only_CC)
plot(simulationOutput)

testDispersion(simulationOutput) # not overdispersed
testZeroInflation(simulationOutput) # not zero inflated
```


#6. P. achatinus exclusive analysis

##a. Priach percentage visualization
- How does abundance, percentage vary across habitat types?
```{r}
# site_labels here includes all 4 site types b/c comparing percentages
site_labels <-  c("Shade Cacao \n (n = 12)", 
                  "Sun Cacao \n (n = 11)", 
                  "Abandoned Plantation \n (n = 4)",
                  "Remnant Forest \n (n = 5)")

# reorder factor levels
priach_by_site$Tipo <- factor(priach_by_site$Tipo,
                         c("N", "C", "V", "B"))

# plot P. achatinus across landuse types using palet, my_theme
plot_priach_by_site <- ggplot(priach_by_site, 
                                     aes(x = Tipo, y = priach_percentage, 
                                         fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Land use type",
       y = "P. achatinus percentage of total captures")

plot_priach_by_site

ggsave("output/plot_priach_by_site.png", width = 10, height = 10)
```


##b. Priach percentage Kruskal-Wallis across landuse types
- use Kruskal-Wallis b/c ANOVA residuals not normally distributed
- there is no statistically sig difference in priach % between landuse types
```{r}
# ANOVA residuals are not normally distributed
######################################################
# create ANOVA
priach_percentage_aov <- aov(priach_percentage ~ Tipo,
  data = priach_by_site
)
# visually check ANOVA assumption of normality of residuals
par(mfrow = c(1, 2)) # combine plots
# histogram
hist(priach_percentage_aov$residuals)
# QQ-plot
qqPlot(priach_percentage_aov$residuals,
  id = FALSE # id = FALSE to remove point identification
)
# shapiro test for normality - ANOVA is NOT NORMALLY DISTRIBUTED
shapiro.test(priach_percentage_aov$residuals)
######################################################

# so instead use Kruskal-Wallis test b/c doesn't require normality assumptions
priach_percentage_kruskal <- kruskal.test(priach_percentage ~ Tipo, 
                                          data = priach_by_site)
priach_percentage_kruskal
# No significant difference in priach percentage between landuse types
```


#7. Analyses without P. achatinus


Quickly run subset of analyses to see if removing priach changes findings
- is priach driving abundance and (lack of) species richness patterns?

comparing richness and abundance of active cacao sites
```{r}
# richness in cacao no priach
test_plot <- ggplot(site_richness_local_no_priach, 
                                     aes(x = Tipo, y = no_species, fill = Tipo)) +
  geom_boxplot() +
 # scale_fill_manual(values = pal) + # order needs to match factor levels above
  #scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Land use type",
       y = "Non priach Species richness")
test_plot

# abundance in cacao no priach
test_2 <- ggplot(site_richness_local_no_priach, 
                                     aes(x = Tipo, y = num_ind, fill = Tipo)) +
  geom_boxplot() +
  my_theme() + 
  labs(x = "Land use type",
       y = "Num non priach ind")
test_2
```

doesn't appear to be any change in biplot
```{r}
# Define specific colors
colors <- c("#DFC27D", "#A6611A")
names(colors) <- levels(as.factor(site_richness_local_no_priach$Tipo))

# Mapping site type abbreviations to full names
site_type_names <- c(C = "Sun cacao", N = "Shade cacao")

# Map colors to points based on Tipo
point_colors <- colors[as.factor(site_richness_local_no_priach$Tipo)]

# Extract PCA components and loadings
pca_scores <- pca_fit$x  # Observation scores
pca_loadings <- pca_fit$rotation  # Variable loadings

# Create the biplot
plot(pca_scores[, 1], pca_scores[, 2], 
     col = point_colors, 
     pch = 19, 
     xlab = "PC1", 
     ylab = "PC2", 
     main = "Customized PCA Biplot",
     xlim = c(min(c(pca_scores[, 1], pca_loadings[, 1])) * 1.2, 
              max(c(pca_scores[, 1], pca_loadings[, 1])) * 1.2),
     ylim = c(min(c(pca_scores[, 2], pca_loadings[, 2])) * 1.2, 
              max(c(pca_scores[, 2], pca_loadings[, 2])) * 1.2))

# Add arrows for variable loadings
arrows(0, 0, pca_loadings[, 1] * max(abs(pca_scores[, 1])), 
       pca_loadings[, 2] * max(abs(pca_scores[, 2])), 
       length = 0.1, col = "black")

# Add variable labels
text(pca_loadings[, 1] * max(abs(pca_scores[, 1])), 
     pca_loadings[, 2] * max(abs(pca_scores[, 2])), 
     labels = rownames(pca_loadings), 
     col = "black", cex = 0.8)

# Add legend
legend("bottomright", 
       legend = c("Shade", "Sun"), 
       col = c("#A6611A", "#DFC27D"), 
       pch = 19)
```

glmer of site richness without priach shows no difference
```{r}
mod_richness_full_no_priach <- glm(no_species ~ site_mean_CC_scaled +
                                    site_mean_veg_scaled +
                                    water_dist_scaled +
                                    forest_cover_400_scaled +
                                    Elevation,
                                    data = site_richness_local_no_priach,
                                    family = poisson())

summary(mod_richness_full_no_priach) # nothing sig - AIC 101.11
vif(mod_richness_full_no_priach) # VIF not concern
check_overdispersion(mod_richness_full_no_priach) # not over dispersed

# assess model fit using dHARMA - no issues!
simulationOutput <- simulateResiduals(fittedModel = mod_richness_full_no_priach)
plot(simulationOutput)

testDispersion(simulationOutput) # not over dispersed
testZeroInflation(simulationOutput) # not zero inflated
############# FULL MODEL - nothing sig - AIC 103.5 ##############
```

