---
title: "ranas_community_scrap"
author: "Holden Jones"
date: "2024-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## pair-wise dissimilarity

Calculate pair-wise dissimilarity (distance) using vegdist
- gower and bray-curtis are good in detecting underlying ecological gradients
```{r}
par(mfrow = c(1, 2))
bray <- vegdist(new_mat, "bray") 
gower <- vegdist(new_mat, "gower")

hist(bray, xlim = range(0.0,1.0))
hist(gower, xlim = range(0.0,1.0))
```

## CONTINUE HERE
dissimilarity analysis is good way to explore variability in community comp
- next steps would be to do some sort of cluster analysis
- see where community associations exist
- but switching gears to look at rarefaction


## nmds


## reorder new_mat rows to allow nmds to draw borders of communities
- goal is to represent info from multiple dimensions into a few, so can visualize
  and interpret.
- so here, present the similarity of communities in 2-dimensional space

this is going to be ugly, but have triple checked against original new_mat
```{r}
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )
# Defined the desired order of rows: Shade, Sun, Abandoned, Forest

# Reorder rows
new_mat <- new_mat[new_order, ]
```



## create nmds plot and draw boundaries of landuse types

code from Peat Clark's vegan tutorial
```{r}
new_mat_NMDS=metaMDS(new_mat, k=2)
# use community by species matrix from above

plot(new_mat_NMDS)
# here sites are open circles, species are red crosses

# here help visualize by labeling specific sites and species

# ordination plot function especially for congested plots
ordiplot(new_mat_NMDS,type="n")
# this function adds text or points to ordination plots
orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

orditorp(new_mat_NMDS,display="sites",cex=1.25,air=0.01)
```

## use ordihull to help visualize different treatments unifying sites

assign treatment levels per landuse type on reordered new_mat_reordered
```{r}
landuse_treatments=c(rep("Shade",12),
                     rep("Sun",11), 
                     rep("Abandoned",4), 
                     rep("Forest",5))
# ensure that labels are in correct order and have correct length

ordiplot(new_mat_NMDS,type="n")

ordihull(new_mat_NMDS,groups=landuse_treatments,draw="polygon",
         col="grey90",label=F)

orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

# here colors need to align with landuse_treatments from above!
# of course need to use chocolate!
orditorp(new_mat_NMDS,display="sites",col=c(rep("chocolate4",12),
                                            rep("chocolate1",11),
                                            rep("lightgreen",4),
                                            rep("darkgreen",5)), 
         air=0.01,cex=1.25)
```

## very cool! see a lot of overlap in species simmilarity across landuse types
- but, forest and abandoned sites are similar, minus 3 outliers (FLL-B, FLL-C, FCAT-B3)
- these are also the lower elevation sites of the 9!
- think that elevation will be an important thing to include when comparing communities!
## CONTINUE HERE looking at elevation and group identity for NMDS

## NEXT STEPS: MAKE ORDINATION OFF GROUP, COMMUNITY DISSIMILARITY





