
# PROCESSING DATA

Use this script to perform small tasks on raw data to make it tidier, 
easier to work with. Output is processed data to be used in analysis scripts.


# load packages
```{r}
library(tidyverse)
library(vegan)
library(labdsv)
```

define file path for processed outputs
```{r}
# this will be consistent for each output in this script
data_dir <- file.path(".", "data/processed_data")
```


# ranas

ranas - individual occurrence data across all sites
```{r}
# read raw ranas data
ranas <- read_csv("data/raw_data/ranas_04_05_24.csv")

# format dates
ranas$sample_date <- as.Date(ranas$Fecha, format = "%d/%m/%y")

# need to change VC-V to VC-N, turns out this site was not actually abandoned!
ranas <- ranas %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"),
         year = substr(sample_date, 1, 4))

# remove '..20' column - not sure why these are being added
ranas <- ranas %>%
  select(-...20)

ranas$Tipo[ranas$Sitio == "VC-N"] <- "N"

# remove any Final_ID with sequence - these are awaiting sequencing results
ranas <- ranas %>% 
  filter(Final_ID != "sequence")

# change all epibou to epiesp 
# Lopez-Hervas et al. 2024 show that epibou in this area is actually epiesp
ranas$Final_ID[ranas$Final_ID == "epibou"] <- "epiesp"

# make processed csv output
filename <- "ranas_processed.csv"
filepath <- file.path(data_dir, filename)

# write df to csv
write_csv(ranas, file = filepath)
```


# new

new - only new individuals from ranas dataset
```{r}
# make new with only new ind
new <- ranas %>%
  filter(Recap == "N")

# output
filename <- "new_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(new, file = filepath)
```


# new_no_priach

new_no_priach - remove all priach to look at underlying diversity + abundance
```{r}
# make new with only new ind
new_no_priach <- new %>%
  filter(Final_ID != "priach")

# output
filename <- "new_no_priach_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(new_no_priach, file = filepath)
```



# new_threatened

new_threatened - only new individuals from sp of conservation concern - listed as DD or higher
```{r}
new_threatened <- new %>%
  filter(Final_ID == 'epiesp' |
         Final_ID == 'lepper' |
         Final_ID == 'oopsyl' |
         Final_ID == 'leubil' |
         Final_ID == 'prilatic'
           )

# output
filename <- "new_threatened_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(new_threatened, file = filepath)
```


# environmental

environmental - environmental data for each sampling event
```{r}
# environmental data - change VC-V to VC-N
environmental <- read_csv("data/raw_data/environment_04_08_23.csv")

environmental$sample_date <- as.Date(environmental$Fecha, format = "%d/%m/%y")

# update VC-N
environmental <- environmental %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"))

environmental$Tipo[environmental$Sitio == "VC-N"] <- "N"

# output
filename <- "environmental_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(environmental, file = filepath)
```


# site_data

site_data - location, canopy cover, and miscellaneous data for each site
```{r}
# site_data - change VC-V to VC-N
site_data <- read_csv("data/raw_data/site_data_04_28_24.csv")

# update VC-N
site_data <- site_data %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"))

site_data$Tipo[site_data$Sitio == "VC-N"] <- "N"

# output
filename <- "site_data_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(site_data, file = filepath)
```

# local_by_trans

For local - not at a quadrant level, but at a transect level
- Lots of local variables were measured at the quadrant level for practicality

So, here summarize (using mean, coefficient of variation)
- canopy cover
- leaf litter coverage
- vegetation
at a transect level:
```{r}
local_by_trans <- site_data %>%
  rowwise() %>%
  mutate(
    tran_mean_CC = mean(c_across(`CC A`:`CC D`), na.rm = TRUE),
    tran_coef_var_CC = sd(c_across(`CC A`:`CC D`), na.rm = TRUE) / tran_mean_CC
  )
```

Need to fill in vegetation and hojarasca values
- measured during first site visit, if site did not change between subs. visits 
  wrote 'same' - this was the case vast majority of time!
  
So now need to fill these columns based on the values of previous visits:
```{r}
environment_fill <- environmental %>%
  group_by(Sitio, Fecha, Transecto, Bloque) %>%
  mutate(Veg_fill = if_else(Veg == 'same', NA_character_, Veg),
         Hoja_fill = if_else(Hoja =='same', NA_character_, Hoja))

environment_fill <- environment_fill %>%
  group_by(Sitio, Transecto, Bloque) %>%
  fill(Veg_fill, .direction = "down") %>%
  fill(Hoja_fill, .direction = "down") %>%
  ungroup()

print(environment_fill)
```

now can calculate transect visit averages for vegetation and hojarasca
```{r}
environment_fill$Veg_fill <- as.numeric(environment_fill$Veg_fill)
environment_fill$Hoja_fill <- as.numeric(environment_fill$Hoja_fill)

environment_fill_by_trans <- environment_fill %>%
  group_by(Sitio, Fecha, Transecto) %>%
  summarize(tran_mean_veg = mean(Veg_fill, na.rm = TRUE),
            tran_mean_hoja = mean(Hoja_fill, na.rm = TRUE)
            )
```

now join this dataframe with local_by_trans
```{r}
local_by_trans <- left_join(environment_fill_by_trans, local_by_trans,
                                 by = c("Sitio", "Transecto"))

local_by_trans <- local_by_trans %>%
  select(Sitio, Fecha, Transecto, 
         tran_mean_veg, tran_mean_hoja, Pair,
         tran_mean_CC, tran_coef_var_CC)

filename <- "local_by_trans_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(local_by_trans, file = filepath)
```




# landscape_by_site

For landscape - not at a transect level, but at a site level
- Landscape variables were measured at site level for practicality

So, here summarize
- percent surrounding forest cover
- proximity to water
- elevation
at a site level:

# site_type

site_type - retains elevation, tipo data
```{r}
site_type <- site_data %>% # create site_type df - include name, elevation
  filter(Transecto == 1) %>% # so just one row per site
  select(Sitio, Tipo, Elevation, Pair)

# reorder factor levels now for downstream plotting
site_type$Tipo <- factor(site_type$Tipo, c("N", "C", "V", "B"))

# make Pair a factor for glms
# MAY NEED TO SET PAIR AS FACTOR
#site_type$Pair <- factor(site_type$Pair)

# here add harvested column for later analyses
site_type <- site_type %>%
  mutate(harvested = case_when(
    Tipo == "N" ~ "Yes",
    Tipo == "C" ~ "Yes",
    Tipo == "B" ~ "No",
    Tipo == "V" ~ "No")
    )

# output
filename <- "site_type_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(site_type, file = filepath)
```


# richness_by_site

richness by site - just number of species
```{r}
# group by site
richness_by_site <- new %>%
  group_by(Sitio) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# join with site_type to carry over tipo, elevation, harvested info
richness_by_site <- richness_by_site %>%
  left_join(site_type, by = "Sitio")

# Group by land_use and calculate species richness
richness_summary <- richness_by_site %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# output
filename <- "richness_by_site_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(richness_by_site, file = filepath)
```


# richness_by_site_no_priach

richness_by_site_no_priach - just number of species w/out priach
```{r}
# group by site
richness_by_site_no_priach <- new_no_priach %>%
  group_by(Sitio) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# join with site_type to carry over tipo, elevation, harvested info
richness_by_site_no_priach <- richness_by_site_no_priach %>%
  left_join(site_type, by = "Sitio")

# Group by land_use and calculate species richness
richness_summary_no_priach <- richness_by_site_no_priach %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# output
filename <- "richness_by_site_no_priach_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(richness_by_site_no_priach, file = filepath)
```







# FIX WHY IS THIS NOT WORKING? NAS REMAIN WHEN I NEED THEM TO BE ZEROS

# threatened_richness_by_site
number of threatened species
```{r}
# group by site
threatened_richness_by_site <- new_threatened %>%
  group_by(Sitio) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# left_join with site_type to get type, elevation data
threatened_richness_by_site <- site_type %>%
  left_join(threatened_richness_by_site, by = "Sitio")

# change NAs to 0 for plotting
threatened_richness_by_site[is.na(threatened_richness_by_site)] <- 0

# Group by land_use and calculate species richness
threatened_richness_summary <- threatened_richness_by_site %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))

# output
filename <- "threatened_richness_by_site_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(threatened_richness_by_site, file = filepath)
```


# abundance_by_site

just number of ind
```{r}
# note this is same as site_abundance above
abundance_by_site <- new %>%
  group_by(Sitio) %>%
  count()

# join with site_type to carry over tipo, elevation, harvested info
abundance_by_site <- abundance_by_site %>%
  left_join(site_type, by = "Sitio")

abundance_by_site$Tipo <- factor(abundance_by_site$Tipo, c("N", "C", "V", "B"))

# output
filename <- "abundance_by_site_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(abundance_by_site, file = filepath)
```


# abundance_by_site_no_priach

just number of non priach ind
```{r}
# note this is same as site_abundance above
abundance_by_site_no_priach <- new_no_priach %>%
  group_by(Sitio) %>%
  count()

# join with site_type to carry over tipo, elevation, harvested info
abundance_by_site_no_priach <- abundance_by_site_no_priach %>%
  left_join(site_type, by = "Sitio")

abundance_by_site_no_priach$Tipo <- factor(abundance_by_site_no_priach$Tipo, 
                                           c("N", "C", "V", "B"))

# output
filename <- "abundance_by_site_no_priach_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(abundance_by_site_no_priach, file = filepath)
```


# diversity_by_site

use 3 column format to get diversity indices for each site using vegan package

make new_mat - intermediate step for diversity indices from vegan
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat <- new_mat[new_order, ]
```

diversity_by_site - each major diversity metric
```{r}
# use diversity function from vegan package
diversity_by_site <- site_type %>%
    mutate(Shannon_Index = diversity(new_mat, index = "shannon")) %>%
    mutate(Simpson_Index = diversity(new_mat, index = "simpson")) %>%
    mutate(Inv_Simpson_Index = diversity(new_mat, index = "invsimpson"))

# output
filename <- "diversity_by_site_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(diversity_by_site, file = filepath)
```


# diversity_by_site_no_priach

same as above except using new_no_priach
```{r}
three_column_no_priach <- new_no_priach %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column_no_priach <- data.frame(three_column_no_priach)

new_mat_no_priach <- matrify(three_column_no_priach)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat_no_priach <- new_mat_no_priach[new_order, ]
```

diversity_by_site_no_priach - each major diversity metric
```{r}
# use diversity function from vegan package
diversity_by_site_no_priach <- site_type %>%
    mutate(Shannon_Index = diversity(new_mat_no_priach, index = "shannon")) %>%
    mutate(Simpson_Index = diversity(new_mat_no_priach, index = "simpson")) %>%
    mutate(Inv_Simpson_Index = diversity(new_mat_no_priach, index = "invsimpson"))

# output
filename <- "diversity_by_site_no_priach_processed.csv"
filepath <- file.path(data_dir, filename)
write_csv(diversity_by_site_no_priach, file = filepath)
```










