---
title: "ranas_local_landscape"
author: "Holden Jones"
date: "2024-07-01"
output: html_document
---

# Build and assess models to understand importance of local and landscape vars

```{r}
library(tidyverse)
```

GLM for effect of transect level, local variables on response variables
- species richness, diversity, community composition, abundance
- Spencer suggested using an occupancy approach - also what Dan Karp suggested!

```{r}
ranas <- read_csv("data/processed_data/ranas_processed.csv")
ranas <- as_tibble(ranas)

local_by_trans <- read_csv("data/processed_data/local_by_trans_processed.csv")
local_by_trans <- as_tibble(local_by_trans)
```

# need to summarize the species in each transect visit
```{r}
trans_richness <- ranas %>%
  group_by(Sitio, Transecto, Fecha) %>%
  summarize(num_species = n_distinct(Final_ID))

trans_richness_local <- left_join(trans_richness, local_by_trans,
                                 by = c("Sitio", "Transecto", "Fecha"))
```

throw together some glms! starting point!
# don't have a random effect here - probably would want to use this? 
- can't run random effects w/ glm: 1 | Sitio
```{r}
model_richness_trans <- glm(num_species ~ 
                              tran_mean_veg + 
                              tran_mean_hoja + 
                              tran_mean_CC,
                            data = trans_richness_local, 
                            family = poisson()
                            )

# Display the model summary
summary(model_richness_trans)
```

# quick figures here! just for first draft of MS
```{r}
# tran_mean_hoja on num_species
ggplot(trans_richness_local, aes(tran_mean_hoja, num_species)) + 
       geom_point() +
  geom_smooth()

# tran_mean_CC on num_species
ggplot(trans_richness_local, aes(tran_mean_CC, num_species)) + 
       geom_point() +
  geom_smooth()

# tran_mean_veg on num_species
ggplot(trans_richness_local, aes(tran_mean_veg, num_species)) + 
       geom_point() +
  geom_smooth()
```





## THIS DOESNT WORK:

# chat gpt says should use a GLMM - so can incorporate 
```{r}
# library(lme4)
# library(Matrix)
# 
# model_richness_trans <- glmer(num_species ~ 
#                                 tran_mean_veg + 
#                                 tran_mean_hoja + 
#                                 tran_mean_CC + 
#                                 (1 | Sitio), 
#                               data = trans_richness_local, 
#                               family = poisson()
#                               )
# 
# 
# summary(model_richness_trans)
# ```




