---
title: "Ranas_compare"
author: "S. Holden Jones"
date: "3/6/24"
output: word_document
---

# COMPARE RICHNESS, ABUNDANCE, DIVERSITY, COMPOSITION ACROSS LANDUSE TYPES


This script will take processed data as input, run statistic tests and produce
figures comparing the dependent variables (diversity, abundance, composition) of
each landuse type broadly. Other scripts will explore the variation in these
response variables w/in specific landuse types.


load packages
```{r, include=FALSE}
library(labdsv)
library(tidyverse)
library(vegan)
library(iNEXT)
library(sjmisc)
library(RColorBrewer) # for assigning plot colors
library(DHARMa) # for assessing model fit
library(SuppDists) # needed for assessing additional dist fits
```


load data
```{r, include=FALSE}
ranas <- read_csv("data/processed_data/ranas_processed.csv")
ranas <- as_tibble(ranas)

new <- read_csv("data/processed_data/new_processed.csv")
new <- as_tibble(new)

new_threatened <- read_csv("data/processed_data/new_threatened_processed.csv")
new_threatened <- as_tibble(new_threatened)

site_data <- read_csv("data/processed_data/site_data_processed.csv")
site_data <- as_tibble(site_data)

site_type <- read_csv("data/processed_data/site_type_processed.csv")
site_type <- as_tibble(site_type)

richness_by_site <- read_csv(
  "data/processed_data/richness_by_site_processed.csv")
richness_by_site <- as_tibble(richness_by_site)

threatened_richness_by_site <- read_csv(
  "data/processed_data/threatened_richness_by_site_processed.csv")
richness_by_site <- as_tibble(threatened_richness_by_site)

abundance_by_site <- read_csv(
  "data/processed_data/abundance_by_site_processed.csv")
abundance_by_site <- as_tibble(abundance_by_site)

diversity_by_site <- read_csv(
  "data/processed_data/diversity_by_site_processed.csv")
diversity_by_site <- as_tibble(diversity_by_site)
```

define a consistent palet, labels, and theme to be used in figures
## still would be nice to have palet for harvested comparison
```{r}
# palet for comparisons of all 4 site types
pal <- brewer.pal(4, "BrBG")

# site labels for 4 site comparison, use in several plots
site_labels <-  c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")

# site labels for harvest, non-harvest comparison
labels_harvest <-  c("Not harvested \n (n = 9)", "Harvested \n (n = 23)")

# set a theme for ggplots
my_theme <- function() {
  theme_minimal() + 
   theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(size = 14, color = "gray25"),
        legend.text = element_text(size = 12))
}
```












# RICHNESS ACROSS LANDUSE TYPES



## 30 total species
```{r}
new %>%
  distinct(Final_ID) %>%
  count()
```


## before selecting statistical test, need to look at dist. of richness by site
- use Poisson dist. b/c using count data with no upper limit!
```{r}
(richness_by_site_hist <- ggplot(richness_by_site, aes(no_species, fill = Tipo)) +
  geom_histogram())
```


## run glm for species richness across landuse types, using Poisson dist.


## NO SIG DIFF IN RICHNESS ACROSS LANDUSE TYPES
- bosque is sig diff from nacional w/out elevation as random fact - viejo almost
- bosque and viejo very close to being significant - need final IDs back!
- using poisson dist in glm - DHARMa simulation shows that it works well
```{r}
richness_by_type_glm <- glm(no_species ~ Tipo + (1|Elevation), 
                            data = richness_by_site, 
             family = poisson(link = "log"))
summary(richness_by_type_glm)

# here use DHARMa package to see if we're using the correct model
# no significant problems detected!
simulationOutput_richness_by_type <- simulateResiduals(
  fittedModel = richness_by_type_glm)
plot(simulationOutput_richness_by_type)
```


plot species richness across landuse types
```{r}
# reorder factor levels
richness_by_site$Tipo <- factor(richness_by_site$Tipo,
                         c("N", "C", "V", "B"))

# plot species richness across landuse types using palet, my_theme
plot_richness_by_site <- ggplot(richness_by_site, 
                                     aes(x = Tipo, y = no_species, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Landuse types",
       y = "Species richness",
       title = "Species richness across landuse types")
plot_richness_by_site
```






# RICHNESS ACROSS HARVESTED UNHARVESTED







## run glm for harvest on species richness using Poisson dist

## SIG HIGHER RICHNESS IN NON_HARVESTED SITES

higher richness in non-harvested sites
- using Poisson dist
- including elevation as random effect in glm
```{r}
richness_by_harvest_glm <- glm(no_species ~ harvested + (1|Elevation), 
                               data = richness_by_site, 
                               family = poisson(link = "log"))
summary(richness_by_harvest_glm)

# within group deviation from uniformity n.s.
# Levene Test for homogeneity of variance n.s.
simulationOutput_richness_by_harvest <- simulateResiduals(
  fittedModel = richness_by_harvest_glm)
plot(simulationOutput_richness_by_harvest)
```


B+V have sig higher richness than N+C
## aesthetically would want to mix colors from previous plot
```{r}
plot_richness_by_harvest <- ggplot(richness_by_site, 
                                   aes(x = harvested, 
                                       y = no_species, fill = harvested)) +
  scale_x_discrete(labels = c(labels_harvest)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Active cacao plantation?", y = "Species richness") +
  ggtitle("Species richness in harvested and non-harvested areas")
plot_richness_by_harvest
```






# RICHNESS OF THREATENED SPECIES ACROSS LANDUSE TYPES






here just focus on species listed by IUCN as NT or higher, or DD
```{r}
threatened_richness_by_site$Tipo <- factor(threatened_richness_by_site$Tipo,
                         c("N", "C", "V", "B"))

# plot species richness across landuse types using palet, my_theme
plot_threatened_richness_by_site <- ggplot(threatened_richness_by_site,
     aes(x = Tipo, y = no_species, fill = Tipo)) +
     geom_boxplot() +
     scale_fill_manual(values = pal) +
   # order needs to match factor levels above
     scale_x_discrete(labels = c(site_labels)) +
     my_theme() +
     labs(x = "Landuse types", y = "Threatened species richness",
     title = "Threatened species richness across landuse types")
plot_threatened_richness_by_site
```











# RAREFIED RICHNESS ACROSS LANDUSE TYPES


need to go through rarefy_scrap to piece this piece together
- use Chao iNEXT package









# ABUNDANCE ACROSS LANDUSE TYPES





here just simply number of new individuals 
- could eventually complicate this to look at pop size estimates?




## before selecting statistical test, need to look at dist. of abundance by site
- use Poisson dist. b/c using count data with no upper limit!
```{r}
(abundance_by_site_hist <- ggplot(abundance_by_site, aes(n, fill = Tipo)) +
  geom_histogram(binwidth = 5))
```



## run glm for abundance across landuse types, using Gaussian dist. Elev rand eff

DONT RUN POISSON HERE!
- quantile deviations detected
- combined adjusted quantile test significant

## NO SIG DIFF IN ABUNDANCE ACROSS LANDUSE TYPES

- elevation random effect
```{r}
# default family dist is Gaussian
abundance_by_type_glm <- glm(n ~ Tipo + (1|Elevation), data = abundance_by_site)
summary(abundance_by_type_glm)

# no significant problems detected w/Gaussian!
simulationOutput_abundance_by_type <- simulateResiduals(
  fittedModel = abundance_by_type_glm)
plot(simulationOutput_abundance_by_type)
```

```{r}
plot_abundance_by_site <- ggplot(abundance_by_site, aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Landuse type",
       y = "Individuals per site",
       title = "Abundance across landuse types")
plot_abundance_by_site
```












# DIVERSITY ACROSS LANDUSE TYPES






Look at how different diversity metrics compare across landuse types


## look at dist of diversity metrics
```{r}
# shannon - clearly not normal dist! 
(shannon_hist <- ggplot(diversity_by_site, aes(Shannon_Index)) +
  geom_histogram(binwidth = .05))

# simpson - clearly not normal dist! 
(simpson_hist <- ggplot(diversity_by_site, aes(Simpson_Index)) +
  geom_histogram(binwidth = .05))

# inv_simpson - clearly not normal dist! 
(inv_simpson_hist <- ggplot(diversity_by_site, aes(Inv_Simpson_Index)) +
  geom_histogram(binwidth = .1))
```

## use Gaussian dist for Shannon, Simpson, inv_Gaussian for Inv_Simpson

Thought Poisson dist would work for these b/c not normal and from count data
- woah! sig deviation for Poisson dist.
- quantile deviations detected
- combined adjusted quantile test sig


## SIG HIGHER SHANNON DIV IN BOSQUE

significantly higher shannon div in Bosque vs. other sites
- viejo is close
```{r}
# shannon w/ Gaussian dist - all good
shannon_by_type_glm <- glm(Shannon_Index ~ Tipo + (1|Elevation), 
                           data = diversity_by_site)
summary(shannon_by_type_glm)

# no significant problems detected using Gaussian dist!
simulationOutput_shannon_by_type <- simulateResiduals(
  fittedModel = shannon_by_type_glm)
plot(simulationOutput_shannon_by_type)
```


## SIG HIGHER SIMPSON DIV IN BOSQUE

sig higher simpson div in Bosque vs. other sites
- viejo not particularly close
```{r}
# simpson w/ Gaussian dist
simpson_by_type_glm <- glm(Simpson_Index ~ Tipo + (1|Elevation), 
                           data = diversity_by_site)
summary(simpson_by_type_glm)

# no significant problems detected using Gaussian dist!
simulationOutput_simpson_by_type <- simulateResiduals(
  fittedModel = simpson_by_type_glm)
plot(simulationOutput_simpson_by_type)
```

## NO SIG DIFF IN INV_SIMPSON DIV ACROSS LANDUSE TYPES

no sig difference in inv_simpson diversity across landtypes
- Bosque .18, viejo not close

NOTE THAT HERE USING INVERSE GAUSSIAN DIST
- Gaussian - quantile deviations detected - comb adjust - quant test sig
- Binomial - error - y values must be b/ween 0:1
- Gamma - deviation significant w/ outlier test but better than others
- Inv Gaussian - works! need to load additional cran package SuppDists
```{r}
# inv_simpson w/ inverse gaussian dist.
inv_simpson_by_type_glm <- glm(Inv_Simpson_Index ~ Tipo + (1|Elevation), 
                               data = diversity_by_site,
                               family = inverse.gaussian(link = "1/mu^2"))
summary(inv_simpson_by_type_glm)

# no significant problems detected using inverse Gaussian dist!
# all other tested dist threw errors w/ DHARMa
simulationOutput_inv_simpson_by_type <- simulateResiduals(
  fittedModel = inv_simpson_by_type_glm)
plot(simulationOutput_inv_simpson_by_type)
```

## plot how diversity indices vary across land types
- these are pretty much identical regardless of index used

shannon
```{r}
plot_shannon <- ggplot(diversity_by_site, 
                       aes(x = Tipo, y = Shannon_Index, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Landtype",
       y = "Shannon diversity",
       title = "Shannon diversity across landuse types")
plot_shannon
```

simpson
```{r}
plot_simpson <- ggplot(diversity_by_site, 
                       aes(x = Tipo, y = Simpson_Index, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Landtype",
       y = "Simpson diversity",
       title = "Simpson diversity across landuse types")
plot_simpson
```

inv simpson
```{r}
plot_inv_simpson <- ggplot(diversity_by_site, 
                           aes(x = Tipo, y = Inv_Simpson_Index, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Ecological landtype",
       y = "Inverse simpson diversity",
       title = "Inverse Simpson diversity across landuse types")
plot_inv_simpson
```











# DIVERSITY ACROSS HARVEST, UNHARVESTED


do the same thing as above, compare each index using harvested, unharvested column
- glms across harvested, unharvested




## use Gaussian dist for Shannon, Simpson, Inv_Simpson
NOTE THAT HERE THE GAUSSIAN DIST WORKS FINE FOR INV_SIMPSON





## SIG HIGHER SHANNON DIV IN NON-HARVESTED SITES

check w/ interpretation of this glm summary
```{r}
# using Gaussian dist.
shannon_harvest_glm <- glm(Shannon_Index ~ harvested + (1|Elevation), 
                           data = diversity_by_site)
summary(shannon_harvest_glm)

# no issues with Gaussian dist
simulationOutput_shannon_harvest <- simulateResiduals(
  fittedModel = shannon_harvest_glm)
plot(simulationOutput_shannon_harvest)
```


## SIG HIGHER SIMPSON DIV IN NON-HARVESTED SITES

check interpretation of this glm summary
```{r}
# using Gaussian dist.
simpson_harvest_glm <- glm(Simpson_Index ~ harvested + (1|Elevation), 
                           data = diversity_by_site)
summary(simpson_harvest_glm)

# no issues with Gaussian dist
simulationOutput_simpson_harvest <- simulateResiduals(
  fittedModel = simpson_harvest_glm)
plot(simulationOutput_simpson_harvest)
```

## NO SIG DIFF IN IN_SIMPSON ACROSS HARVEST, NON-HARVEST

check interpretation of this glm summary
```{r}
# using Gaussian dist.
inv_simpson_harvest_glm <- glm(Inv_Simpson_Index ~ harvested + (1|Elevation), 
                               data = diversity_by_site)
summary(inv_simpson_harvest_glm)

# no issues with Gaussian dist
simulationOutput_inv_simpson_harvest <- simulateResiduals(
  fittedModel = inv_simpson_harvest_glm)
plot(simulationOutput_inv_simpson_harvest)
```

## make figures for each
- aesthetically think about fill colors

shannon_harvest
```{r}
plot_shannon_harvest <- ggplot(diversity_by_site, 
                               aes(x = harvested, y = Shannon_Index, 
                                   fill = harvested)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Harvested",
       y = "Shannon diversity",
       title = "Shannon diversity across harvested and unharvested sites")
plot_shannon_harvest
```

simpson_harvest
```{r}
plot_simpson_harvest <- ggplot(diversity_by_site, 
                               aes(x = harvested, y = Simpson_Index, 
                                   fill = harvested)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Harvested",
       y = "Simpson diversity",
       title = "Simpson diversity across harvested and unharvested sites")
plot_simpson_harvest
```

inv_simpson_harvest
```{r}
plot_inv_simpson_harvest <- ggplot(diversity_by_site, 
                               aes(x = harvested, y = Inv_Simpson_Index, 
                                   fill = harvested)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Harvested",
       y = "Inverse Simpson diversity",
       title = "Inverse Simpson diversity across harvested and unharvested sites")
plot_inv_simpson_harvest
```

















# COMPOSITION ACROSS LANDUSE TYPES - different script?


Variety of ways to look at this, can't do all of them!
- NMDS - more for visualizing differences - start here
- Bray-Curtiss Dissimilarity Index - quantifies dissimilarity - 0:1 (same:diff)
  - this one preferred b/c incorporates species abundances
- Jaccard Similarity Index - doesn't incorporate species abundances
- Analysis of similarities for statistics


## NEED TO GET THIS CODE REVIEWED - NOT SURE IF CORRECT


## create new_mat matrix

new_mat - three column format to use in labdsv package
        - reorder rows to allow NMDS to draw by site type
        - creating here b/c was running into row names issue importing from
          processing script
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)

# Define the desired order of rows: Shade, Sun, Abandoned, Forest
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )

# Reorder rows
new_mat <- new_mat[new_order, ]
```


## create nmds plot
-  draw boundaries of landuse types
```{r}
# assign treatment levels per landuse type on reordered new_mat
landuse_treatments=c(rep("Shade",12),
                      rep("Sun",11), 
                      rep("Abandoned",4), 
                      rep("Forest",5))

# use community by species matrix - stress = 0.26 - okay representation
new_mat_NMDS = metaMDS(new_mat, k = 2)

# sites are open circles, species are red crosses
plot(new_mat_NMDS)

# help visualize by labeling specific sites and species

# ordination plot function especially for congested plots
ordiplot(new_mat_NMDS,type="n")
# this function adds text or points to ordination plots
orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

orditorp(new_mat_NMDS,display="sites",cex=1.25,air=0.01)

ordiplot(new_mat_NMDS,type="n")

ordihull(new_mat_NMDS,groups=landuse_treatments,draw="polygon",
         col="grey90",label=F)

orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

# here colors need to align with landuse_treatments from above!
# of course need to use chocolate!
orditorp(new_mat_NMDS,display="sites",col=c(rep("chocolate4",12),
                                            rep("chocolate1",11),
                                            rep("lightgreen",4),
                                            rep("darkgreen",5)), 
         air=0.01,cex=1.25)
```







# INCORPORATE iNEXT PACKAGE

not sure how to interpret this data


## look at rarefaction curves of all sites at once

all sites - pretty crazy
```{r}
# just forest sites:

# okay, so it seems like it thinks that species are sites and sites are species
# can we rotate the dataframe?
new_mat_test <- new_mat %>% 
  rotate_df()


out <- iNEXT(new_mat_test, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# Sample-size-based R/E curves, separating plots by "Assemblage"
ggiNEXT(out, type=1, facet.var="Assemblage")
```


## unsightly! too many sites at once - go one site type at a time

try subsetting new_mat into only forest sites for example - may be easier to visualize
```{r}
# just forest sites:
forest_mat <- new_mat[28:32,]

# okay, so it seems like it thinks that species are sites and sites are species
# can we rotate the dataframe?
forest_mat <- forest_mat %>% 
  rotate_df()


out <- iNEXT(forest_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# Sample-size-based R/E curves, separating plots by "Assemblage"
ggiNEXT(out, type=1, facet.var="Assemblage")
```

how about just abandoned sites?
```{r}
# just forest sites:
abandoned_mat <- new_mat[24:27,]

# okay, so it seems like it thinks that species are sites and sites are species
# can we rotate the dataframe?
abandoned_mat <- abandoned_mat %>% 
  rotate_df()


out <- iNEXT(abandoned_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# Sample-size-based R/E curves, separating plots by "Assemblage"
ggiNEXT(out, type=1, facet.var="Assemblage")
```

sun sites
```{r}
# just forest sites:
sun_mat <- new_mat[13:23,]

# okay, so it seems like it thinks that species are sites and sites are species
# can we rotate the dataframe?
sun_mat <- sun_mat %>% 
  rotate_df()


out <- iNEXT(sun_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# Sample-size-based R/E curves, separating plots by "Assemblage"
ggiNEXT(out, type=1, facet.var="Assemblage")
```

shade sites
```{r}
# just forest sites:
shade_mat <- new_mat[1:12,]

# okay, so it seems like it thinks that species are sites and sites are species
# can we rotate the dataframe?
shade_mat <- shade_mat %>% 
  rotate_df()

out <- iNEXT(shade_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# Sample-size-based R/E curves, separating plots by "Assemblage"
ggiNEXT(out, type=1, facet.var="Assemblage")
```




## How about visualize by groups of sites? That could be cool!
- Need to just make different matrices

Grupo Estacion
```{r}
# just estacion sites: order to help with visualizations
rows <- c('LL-C', 'JZ-C', 'JZJ-C', # sun
          'LL-N', 'JZ-N', 'AL-N', # shade
          'FCAT-V1', 'FCAT-V2', # abandoned
          'FCAT-B1', 'FCAT-B2') # forest

estacion_mat <- new_mat[rows,] %>%
  rotate_df()

out <- iNEXT(estacion_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)

ggiNEXT(out, type=1, facet.var="Assemblage")
```

Grupo Laguna
```{r}
# just laguna sites: order to help with visualizations
rows <- c('FLL-C', 'DB-C', # sun
          'JS-N', 'DB-N',  # shade
          'FLL-V',         # abandoned
          'FLL-B')         # forest

laguna_mat <- new_mat[rows,] %>%
  rotate_df()

out <- iNEXT(laguna_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)

ggiNEXT(out, type=1, facet.var="Assemblage")
```

Grupo Gavilanes
```{r}
# just gavilanes sites: order to help with visualizations
rows <- c('DG-C',  # sun
          'DG-N',  # shade
          'DG-V',  # abandoned
          'DG-B')  # forest

gavilanes_mat <- new_mat[rows,] %>%
  rotate_df()

out <- iNEXT(gavilanes_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)

ggiNEXT(out, type=1, facet.var="Assemblage")
```

Grupo Aniseto-Caucho
## note that VC-C and GP-N had extremely low sampling coverage 
- try without these sites?
```{r}
# just aniseto_caucho sites: order to help with visualizations
rows <- c('AZ-C', 'LG-C', 'GP-C', 'PM-C', 'VC-C', # sun
          'AZ-N', 'LG-N', 'GP-N', 'PM-N', 'VC-N', 'PV-N', # shade
          'FCAT-B3') # forest

aniseto_mat <- new_mat[rows,] %>%
  rotate_df()

out <- iNEXT(aniseto_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)

ggiNEXT(out, type=1, facet.var="Assemblage")
```

Grupo Aniseto-Caucho, no VC-C, no GP-N
## note that VC-C and GP-N had extremely low sampling coverage 
- try without these sites?
```{r}
# just aniseto_caucho sites: order to help with visualizations
rows <- c('AZ-C', 'LG-C', 'GP-C', 'PM-C', #'VC-C', # sun
          'AZ-N', 'LG-N', #'GP-N', 
          'PM-N', 'VC-N', 'PV-N', # shade
          'FCAT-B3') # forest

aniseto_mat <- new_mat[rows,] %>%
  rotate_df()

out <- iNEXT(aniseto_mat, q=c(0, 1, 2), datatype="abundance", endpoint=500)

ggiNEXT(out, type=1, facet.var="Assemblage")
```



















# BELOW HERE NOT WORKING - SCRATCH


# WARNING MESSAGES: WARNING THIS SITE HAS ONLY ONE SPECIES






```{r}
# # maybe have to transpose (swap matrix?)
# 
# test <- t(new_mat)
# out <- iNEXT(test, q = 0, datatype = "abundance")
# ggiNEXT(out, type=1, facet.var="Assemblage")
# 
# # str(new_mat)
# # out <- iNEXT(new_mat, q = 0, datatype = "abundance")
# 
# 
# ggiNEXT(out, type=1, facet.var="Assemblage")
# 
# sample_matrix <- matrix(1:9, nrow = 3, ncol = 3, byrow = TRUE)
# 
# # Original matrix
# print("Original matrix:")
# print(sample_matrix)
# 
# # Transpose the matrix using dplyr
# transposed_matrix <- t(sample_matrix)
# 
# # Transposed matrix
# print("Transposed matrix:")
# print(transposed_matrix)
```









## try again here - what is output of bird data?
```{r}
# data(spider)
# out <- iNEXT(spider, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# # Sample-size-based R/E curves, separating plots by "Assemblage"
# ggiNEXT(out, type=1, facet.var="Assemblage")
# 
# 
# test <- as.matrix(new_mat)
# 
# 
# 
# out <- iNEXT(test, q=c(0, 1, 2), datatype="abundance", endpoint=500)
# # Sample-size-based R/E curves, separating plots by "Assemblage"
# ggiNEXT(out, type=1, facet.var="Assemblage")
# 
# 
# data(new_mat)
```







# tutorial from Francisco - need to import matrix as .txt file
# move this to process eventually
https://www.youtube.com/watch?v=4yUXG28QPLM&t=28s
```{r}
# # # Define the file path where you want to save the matrix
# # file_path <- "matrix_data.txt"
# # 
# # # Export the matrix to a text file
# # write.table(matrix_data, file = file_path, sep = "\t", row.names = FALSE, col.names = FALSE)
# # 
# 
# # USING THIS BELOW:
# 
# 
# # this will be consistent for each output in this script
# data_dir <- file.path(".", "data/processed_data")
# 
# # output
# filename <- "test_table_iNEXT.txt"
# filepath <- file.path(data_dir, filename)
# write.table(test, file = filepath)
# ```
# 
# 
# ## my attempt using Francisco's code
# 
# ```{r}
# data.ants <- read.table("data/processed_data/test_table_iNEXT.txt", 
#                            header = TRUE, sep = "\t")
# 
# 
# 
# 
# ```
# 
# 
# 
# 
# 
# 
# 
# ## code from francisco - not working!
# 
# ```{r}
# library(iNEXT)
# library(ggplot2)
# library(devtools)
# library(httr)
# 
# #https://www.youtube.com/watch?v=24PtDhIYg8g tutorial 
# 
# #Step 1 load the data
# 
# iNEXT_matrix = read.table("test_table_iNEXT.txt", head = T)
# 
# head(data.ants)
# 
# #You can upload more matrix to do one 
# 
# #Convert the data frame into a list matrix for inext 
# 
# divh <- as.matrix(apply(data.ants[,-1],2, as.integer))
# 
# #Name as row names
# 
# row.names(divh) <- data.ants[,1]
# 
# #List the matrix or matrices 
# 
# divant = list(Hymenoptera=divh)
# 
# str(divant)
# 
# edit = edit(divant) #to look if all are good 
# 
# #Alpha diversity  
# 
# out.inc = iNEXT(divant, q=0, datatype = "abundance") 
# out.inc
# 
# #SC sampling completeness 
# 
# #For ants: 84.62% (n of 65, ci= 78-90) duplication of the sample effort mean into 93.42% of 
# #the SC 
# 
# #Figures 
# 
# #FigDivAlpha = ggiNEXT(out.inc, type = 1, color.var <- "Order.q", z <- fortify(out.inc), 
#                      # z$col <- z$shape <- factor(z$Order.q)) Revisar 
# par(mfrow=c(2,3))
# 
# #figures species diversity 
# fig1 <- plot(out.inc, col= "Blue")
# fig1
# #figures sample coverage  
# fig2 <-plot(out.inc, type = 2, col= "Blue")
# fig2
# #figures species diversity against sample coverage   
# fig3 <-plot(out.inc, type = 3, col= "Blue")
# fig3
# 

```








 
# ERRORS SHOWING UP BELOW HERE


## calculate Bray-Curtis Index
- maybe move this to ranas_process actually! 
```{r}
# # Here calculating distance first, then creating NMDS off this
# dist <- vegdist(new_mat, method="bray")
# 
# # set seed so consistent across iterations of nmds
# set.seed(19960710)
# 
# nmds <- metaMDS(dist)
# # very low stress!
# 
# scores(nmds) %>%
#   as_tibble(rownames = "Group") %>% # needs to be converted to tibble for ggplot2
#   ggplot(aes(x = NMDS1, y = NMDS2)) +
#   geom_point()
```


## can add rarefaction into vegdist!
```{r}
# # need to set sanmple here, what should it be? randomly picked 50
# dist <- avgdist(new_mat, dmethod = "bray", sample = 50)
# # this is a rarefied dist - trying to account for uneven sampling effort
# # these metrics are very sensitive to uneven sampling effort
# 
# nmds <- metaMDS(dist)
# 
# scores(nmds) %>%
#   as_tibble(rownames = "Group") %>%
#   ggplot(aes(x = NMDS1, y = NMDS2)) +
#   geom_point()
```









## WHY ARENT ROW NAMES BEING USED? HEATMAP UNINTELLIGBLE

```{r}
# Step 1: Calculate Bray-Curtis distance
# bray_matrix <- vegdist(new_mat, method="bray")
# 
# # Step 2: Convert distance matrix to percentage similarity
# # Convert distances to similarities (1 - distance)
# similarity_matrix <- 1 - as.matrix(bray_matrix)
# 
# # Step 3: Plot the percentage similarity matrix
# # Example heatmap plot
# heatmap(similarity_matrix, 
#         main="Percentage Similarity Heatmap",
#         xlab="Samples", ylab="Samples",
#         labRow=NA, labCol=NA,
#         col=rev(heat.colors(10)))  # Reversed heat colors for better visualization
```




## adjacent, random piece from Riffomonas project





```{r}
# # ordination of bray_matrix using cmdscale
# 
# cmdscale(bray_matrix) %>% plot()
# 
# pcoa <- cmdscale(bray_matrix)
# colnames(pcoa) <- c("pcoa1", "pcoa2")
# pcoa %>% head()
```




## adjacent, jaccard similarity index






```{r}
# # create a sample data frame with mixed data types 
# df <- data.frame( 
#   numeric_var = c(1,2,3,4), 
#   categorical_var = c("A", "B", "A", "C"), 
#   binary_var = c(0,1,0,1) 
# ) 
#   
# # convert categorical variable to binary 
# df$categorical_var <- ifelse(df$categorical_var == "A", 1, 0) 
#   
# # convert binary variable to binary 
# df$binary_var <- as.numeric(df$binary_var) 
#   
# # calculate the Jaccard distance  
# # matrix using the dist function 
# jaccard_distance_matrix <- dist(df, 
#                                 method = "binary") 
#   
# #Print the distance matrix 
# print(jaccard_distance_matrix)
```




