Use this markdown to go in depth in diversity analyses in cacao_ranas data
- Simpson's - closer to 1 = higher diversity
- Shannon-Weiner - typically from 1.5 - 3.4, higher = more diverse
- Evenness


## NEED TO REVIEW WHOLE SCRIPT - with final data ##

--------------------------------------------------------------------------------
# Load packages, data

use labdsv to get cacao data frame into community data matrix, as needed by
vegan package

load packages
```{r}
library(labdsv)
library(tidyverse)
library(vegan)
library(RColorBrewer)
```

load data, create new df
```{r}
# ranas data
ranas <- read_csv("data/ranas_05_17_23.csv")
ranas$sample_date <- as.Date(ranas$Fecha, format = "%d/%m/%y")

## need to change VC-V to VC-N, turns out this site is not actually abandoned
ranas <- ranas %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N"),
         year = substr(sample_date, 1, 4))
ranas$Tipo[ranas$Sitio == "VC-N"] <- "N"

# make new with only new ind
new <- ranas %>%
  filter(Recap == "N")

# environmental data - change VC-V to VC-N
environmental <- read_csv("data/environment_04_08_23.csv")
environmental$sample_date <- as.Date(environmental$Fecha, format = "%d/%m/%y")

environmental <- environmental %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N")) 
environmental$Tipo[environmental$Sitio == "VC-N"] <- "N"

# site_data - change VC-V to VC-N
site_data <- read_csv("data/site_data.csv")

site_data <- site_data %>%
  mutate(Sitio = replace(Sitio, Sitio == "VC-V", "VC-N")) 
site_data$Tipo[site_data$Sitio == "VC-N"] <- "N"
```


--------------------------------------------------------------------------------
# Simple richness by type

richness by type - just number of species
##NEED to control for sampling effort - ie divide by number of sites##
```{r}
richness_by_type <- ranas %>%
  group_by(Tipo) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

richness_by_type %>%
  ggplot(aes(Tipo, no_species)) +
  geom_col()
```

--------------------------------------------------------------------------------
# Aov for species richness by land type

https://rpubs.com/an-bui/vegan-cheat-sheet

There is no sig difference in sp richness b/ween landtypes
```{r}
sppr <- specnumber(new_mat)

sppr_aov <- aov(sppr ~ Tipo, data = site_type)
summary(sppr_aov)
# p-value of 0.805! no sig difference in sp richness b/ween landtypes
```

## no sig difference in sp richness b/ween landtypes!

Plot species richness b/ween landtypes
```{r}
sppr_df <- sppr %>% 
  enframe() %>% 
  full_join(site_type, by = c("name" = "Sitio"))

sppr_df$Tipo <- factor(sppr_df$Tipo,     # Reorder factor levels
                         c("B", "V", "N", "C"))

pal <- brewer.pal(4, "BrBG")

plot_sppr <- ggplot(sppr_df, aes(x = Tipo, y = value, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")
    ) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Ecological landtype",
       y = "Confirmed species per site",
       title = "Species richness")
plot_sppr
```


--------------------------------------------------------------------------------
# 3 column format

turn ranas into 3 column format - habitat, species, abundance for labdsv
- use matrify!

new_mat - three column format to use in labdsv package
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)
```


--------------------------------------------------------------------------------
# Diversity Indices

Wonderful tutorials!
https://peat-clark.github.io/BIO381/veganTutorial.html
https://www.youtube.com/watch?v=wq1SXGQYgCs

use 3 column format to get diversity indices for each site!
```{r}
print(shannon <- diversity(new_mat, index = "shannon"))
print(simpson <- diversity(new_mat, index = "simpson"))
print(inv_simpson <- diversity(new_mat, index = "invsimpson"))
```

compare these indices with histograms
```{r}
par(mfrow = c(1, 2))  # use par to generate panels with 1 row of 2 graphs
hist(simpson)
hist(shannon)
```

## no sig differences for diversity indices by land type
```{r}
site_type <- site_data %>%
  filter(Transecto == 1) %>% # so just one row per site
  select(Sitio, Tipo)

shannon_aov <- aov(shannon ~ Tipo, data = site_type)
summary(shannon_aov)
# not sig - p > 0.7!

simpson_aov <- aov(simpson ~ Tipo, data = site_type)
summary(simpson_aov)
# not sig - p > 0.7!

inv_simpson_aov <- aov(inv_simpson ~ Tipo, data = site_type)
summary(inv_simpson_aov)
# not sig - p > 0.7!
```

## plot how diversity indices vary across land types
```{r shannon diversity plot}
shannon_df <- shannon %>% 
  enframe() %>% 
  full_join(site_type, by = c("name" = "Sitio"))

shannon_df$Tipo <- factor(shannon_df$Tipo,     # Reorder factor levels
                         c("N", "C", "V", "B"))

pal <- brewer.pal(4, "BrBG")


plot_shannon <- ggplot(shannon_df, aes(x = Tipo, y = value, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")
    ) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(size = 14, color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Landtype",
       y = "Shannon diversity",
       title = "No significant difference in Shannon diversity across landtype 
       (p 0.758)")
plot_shannon
```

```{r simpson diversity plot}
simpson_df <- simpson %>% 
  enframe() %>% 
  full_join(site_type, by = c("name" = "Sitio"))

simpson_df$Tipo <- factor(simpson_df$Tipo,     # Reorder factor levels
                         c("N", "C", "V", "B"))

plot_simpson <- ggplot(simpson_df, aes(x = Tipo, y = value, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")
    ) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(size = 14, color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Landtype",
       y = "Simpson diversity",
       title = "There is no significant difference (p = 0.715) in Simpson diversity across landtypes")
plot_simpson
```

```{r inv simpson diversity plot}
inv_simpson_df <- inv_simpson %>% 
  enframe() %>% 
  full_join(site_type, by = c("name" = "Sitio"))

inv_simpson_df$Tipo <- factor(inv_simpson_df$Tipo,     # Reorder factor levels
                         c("N", "C", "V", "B"))

plot_inv_simpson <- ggplot(inv_simpson_df, aes(x = Tipo, y = value, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")
    ) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Ecological landtype",
       y = "Inverse simpson diversity",
       title = "There is no significant difference (p = 0.75) in Inverse Simpson diversity across landtypes")
plot_inv_simpson
```


--------------------------------------------------------------------------------
# Pair-wise dissimilarity

Calculate pair-wise dissimilarity (distance) using vegdist
- gower and bray-curtis are good in detecting underlying ecological gradients
```{r}
par(mfrow = c(1, 2))
bray <- vegdist(new_mat, "bray") 
gower <- vegdist(new_mat, "gower")

hist(bray, xlim = range(0.0,1.0))
hist(gower, xlim = range(0.0,1.0))
```

dissimilarity analysis is good way to explore variability in community comp
- next steps would be to do some sort of cluster analysis
- see where community associations exist
- but switching gears to look at rarefaction

--------------------------------------------------------------------------------
# Rarefaction

technique to assess expected species richness
- allows calculation of species richness for given number of samples based on
  construction of rarefaction curves
  
issue is that the larger number of ind sampled, more species will be found
- rarefaction curves created by randomly re-sampling pool of N samples multiple
  times and then plotting average number of species found in each sample
- rarefaction generates expected number of species in small collection of n ind
  drawn from large pool of N samples
- typically grow rapidly at first, then slowly as only rarest species remain to
  be sampled

use rarefy and rarecurve functions
```{r}
spAbund <- rowSums(new_mat)
spAbund

test <- colSums(new_mat)
test
```

rarefaction uses smallest number of obs per sample to extrapolate expected
number if all other samples only had that number of obs
- VC-C only had 3!!! may want to proceed without this  in the future
- GP-N also only 9, without these two the next lowest is 22
```{r}
# including VC-C raremin is 3
raremin <- min(rowSums(new_mat))
raremin

sRare <- rarefy(new_mat, raremin)
sRare

# it won't allow us to override this to 22 (makes sense)
# so will need to remove VC-C, GP-N from sites, via row position
new_mat_abundant <- new_mat %>%
  filter(!row_number() %in% c(19, 31))

# let's see what happens with this as our raremin, and only using 30/32 sites
raremin_abundant <- min(rowSums(new_mat_abundant))
raremin_abundant

# sRare gives expected 'rarefied' number of species (not obs) if only 22 collected
sRare_abundant <- rarefy(new_mat_abundant, raremin_abundant)
sRare_abundant

shannon_aov <- aov(shannon ~ Tipo, data = site_type)
summary(shannon_aov)
```

## no statistical difference in rarefied species richness across types
*note that using rarefaction to remove effects of diff sample sizes is bad!*
```{r}
# join tipo data with sRare_abundant in a df
sRare_abundant_df <- sRare_abundant %>% 
  enframe() %>% 
  full_join(site_type, by = c("name" = "Sitio"))

sRare_abundant_aov <- aov(value ~ Tipo, data = sRare_abundant_df)
summary(sRare_abundant_aov)
# p-value is 0.183 (04/9/23) - check back in when sequencing is done!
```

visualization using rarecurve, for ggplot?
## make rarecurve for each type of site
## stuck!
```{r}
# rarecurve for each site, would be nice to have for each type!
rarecurve(new_mat_abundant, col = "blue", cex = 0.4)

# TEST - try to make separate curve for each type
# make dfs of equal length
#site_type_abundant <- site_type %>%
#                      filter(!Sitio == 'VC-C',
#                             !Sitio == 'GP-N')


# may need these as vector eventually?
#test <- cbind(new_mat_abundant, site_type_abundant)

#rarecurve_test <- rarecurve(test, step=1, label=TRUE, col = test$Tipo, 
#                            xlab = "Number of individuals sampled", ylab = "Species richness")
```

boxplot of rarefied species richness
```{r}
# can compare rarefied species richness by type
sRare_abundant_df$Tipo <- factor(sRare_abundant_df$Tipo,
                         c("N", "C", "V", "B"))

plot_sRare_abundant <- ggplot(sRare_abundant_df, aes(x = Tipo, y = value, 
                                                     fill = Tipo)) +
geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 11)", "Sun \n (n = 10)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")
    ) +
        # updated labels to reflect n = 10 for C and N sites
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(size = 14, color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Landtype",
       y = "Rarefied species richness per site",
       title = "There is no significant difference (p = 0.183) in rarefied species richness across landtypes")
plot_sRare_abundant
```


--------------------------------------------------------------------------------
# Non-Metric Multidimensional Scaling

explore some ordination techniques
- PCA
- NMDS
- Correspondence Analysis (CA)

vegan is especially good at NMDS
- goal of NMDS is to collapse info from multiple dimensions into just a few
  so they can be visualized and interpreted.
- NMDS does not produce statistical output - although we could do so

*goal of NMDS is to represent position of communities in multidim space*
*accurately as possible using less dims that can be easily visualized*

NMDS does not use absolute abundances of sp in comm, but rather rank orders
- thus is a felxible technique that accepts variety of data types
- where non-metric comes from

to run NMDS use function metaMDS
```{r}
new_big_NMDS <- metaMDS(new_mat_abundant, k = 2)
```

stress rule of thumb:
- > 0.05 is excellent
- > 0.1 is great
- > 0.2 is good/okay
- > 0.3 is poor

now plot the NMDS
```{r}
plot(new_big_NMDS)

# this shows us both communities (sites, open circles) and species(red crosses)

# use ordiplot and orditorp to add text to plot instead of points
ordiplot(new_big_NMDS,type="n")
orditorp(new_big_NMDS,display="species",col="red",air=0.01) #The function adds text or points to ordination plots
orditorp(new_big_NMDS,display="sites",cex=1.25,air=0.01)
```

## Stuck!
*how to visualize groups / treatments in NMDS?*

using ordihull can draw convex hulls connecting vertices of points from same / 
  different groups
*unsure how to assign each site into its proper treatment or group*


--------------------------------------------------------------------------------
# Abundance variation between site types

plot abundance by type
```{r plot abundance by type}
new_count <- new %>%
  group_by(Sitio) %>%
  count() %>%
  full_join(site_type, by = "Sitio")

new_count$Tipo <- factor(new_count$Tipo, c("N", "C", "V", "B"))

plot_new_count <- ggplot(new_count, aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(size = 14, color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Landtype",
       y = "Individuals per site",
       title = "No significant difference (p = 0.715) in abundance across landtypes")
plot_new_count
```

anova for abundance
## abundance differences are not statistically significant
```{r}
abundance_aov <- aov(n ~ Tipo, data = new_count)
summary(abundance_aov)
```

curious - what if we hadn't done mark recapture, what would abundance look like?
*no difference - still don't see change in abundance*
```{r}
ranas_count <- ranas %>%
  group_by(Sitio) %>%
  count() %>%
  full_join(site_type, by = "Sitio")

ranas_count$Tipo <- factor(ranas_count$Tipo, c("N", "C", "V", "B"))

# quick anova
ranas_abundance_aov <- aov(n ~ Tipo, data = ranas_count)
summary(ranas_abundance_aov)

# and visualization
plot_ranas_count <- ggplot(ranas_count, aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Ecological landtype",
       y = "Observed individuals per site",
       title = "Abundance w/out mark recapture")
plot_ranas_count
```


--------------------------------------------------------------------------------
# Which nested variables to retain?

Talking with Miriam, one of the first things to do is figure out which variables
I need to keep at a bloque, transect, and site level. Start at the smallest level
- I recorded canopy coverage, veg coverage, leaf coverage at bloque level
- Hora, temp, clima, humedad at transect level
- Tipo, Bosque, Pesticidia, Dueno, Grupo, Location at site level

Let's start by averaging all bloque-level variables to a transect level
- start with those in environmental df - veg, hoja
```{r make environmental_trans df}
# in order to calculate mean need to remove same, replace with NAs
# this is fine as all NAs here will mean that cover was same as previous
environmental_trans <- environmental %>%
  mutate(Hoja = na_if(Hoja, "same"),
         Veg = na_if(Veg, "same")) %>%
  mutate_at(c('Hoja', 'Veg', 'Humedad'), as.numeric) %>%
  group_by(Sitio, Fecha, Transecto) %>%
  mutate(transect_hoja = mean(Hoja),
         transect_veg = mean(Veg))

# then can clean up this df, remove unnecessary rows, columns
environmental_trans <- environmental_trans %>%
  filter(Bloque == 'A') %>%
  select(-c('Bloque', 'Hoja', 'Veg'))

# fill these NAs with previous values, use fill function
environmental_trans <- environmental_trans %>% 
  group_by(Sitio, Transecto) %>% 
  fill(transect_hoja, transect_veg)
```

## Distribution of transect_hoja, veg
then look at distribution of transect_hoja, transect_veg
```{r}
# this is interesting! see that viejo and bosque sites have higher hojarasca
# viejo and bosque more consistent intermediate veg, c and n more variation!
ggplot(environmental_trans, aes(transect_hoja, fill = Tipo)) +
  geom_histogram()
ggplot(environmental_trans, aes(transect_veg, fill = Tipo)) +
  geom_histogram()
```

Look at distribution of other transect level variables
```{r}
ggplot(environmental_trans, aes(Hora, fill = Tipo)) +
  geom_histogram()
ggplot(environmental_trans, aes(Temperatura, fill = Tipo)) +
  geom_histogram()
ggplot(environmental_trans, aes(Humedad, fill = Tipo)) +
  geom_histogram()
```

do same thing with transect level site_data
- first will need to pivot_longer to put CCA, CCB, CCC, CCD columns into rows
```{r}
site_data <- site_data %>%
  pivot_longer(
    cols =starts_with("CC"),
    names_to = "Bloque",
    values_to = "Canopy_cover"
  )

# now have column with Bloque, can use this in same code above
# doesn't matter that input is CC A, CC B etc b/c going to delete after :)
site_trans <- site_data %>%
  mutate_at(c('Transecto', 'Canopy_cover'), as.numeric) %>%
  group_by(Sitio, Transecto) %>% # no Fecha b/c only measured once
  mutate(transect_canopy = mean(Canopy_cover))

site_trans <- site_trans %>%
  filter(Bloque == 'CC A') %>%
  select(-c('Bloque', 'Canopy_cover'))
```

## Distribution of canopy cover
histogram looking at distribution of canopy cover across different types
```{r}
ggplot(site_trans, aes(transect_canopy, fill = Tipo)) +
  geom_histogram()
```

------------------------------------------------------------------------------
# Look at distribution of achatinus across active cacao sites
- How many ind can we use from each N and C group for first round of pesticide
- David's group will run 600 swabs for us this summer - how many from each site?
- Power analysis elsewhere shows need 150 from each group to detect 60-40 
  difference between plantation types - do we have enough for this?
- Roughly 15 from each season for each of 20 cacao sites (10 nacional, 10 ccn)
```{r}
# make achatinus subset df here
achatinus <- new %>%
  filter(Final_ID == "priach")

# group by each site - how many in each?
achatinus %>%
  filter(Tipo == "N"| # only planted cacao sites
         Tipo == "C") %>%
  group_by(Sitio) %>%
  summarise(n = n()) %>%
  filter(n > 10) # will need at least 20 so filter these

# how about for each site in each year?
achatinus %>%
  filter(Tipo == "N"|
         Tipo == "C") %>%
  group_by(Sitio, year) %>%
  summarise(n = n()) %>%
  filter(n > 20)

# all sites >= 15
achatinus %>%
  group_by(Sitio, year) %>%
  summarise(n = n()) %>%
  filter(n >= 15)
```


