---
title: "Ranas_compare"
author: "S. Holden Jones"
date: "3/6/24"
output: word_document
---




# COMPARE RICHNESS, ABUNDANCE, DIVERSITY, COMPOSITION ACROSS LANDUSE TYPES














This script will take processed data as input, run statistic tests and produce
figures comparing the dependent variables (diversity, abundance, composition) of
each landuse type broadly. Other scripts will explore the variation in these
response variables w/in specific landuse types.


load packages
```{r, include=FALSE}
library(labdsv)
library(tidyverse)
library(vegan)
library(RColorBrewer) # for assigning plot colors
library(DHARMa) # for assessing model fit
library(SuppDists) # needed for assessing additional dist fits
```

load data
## weird that dfs are of different lengths, why are columns getting added randomly?. Also note that ranas is a tbl_df and new is spec_tbl_df despite being created from ranas

```{r, include=FALSE}
ranas <- read_csv("data/processed_data/ranas_processed.csv")

# remove '...1' column - not sure why this is being added
ranas <- ranas %>%
  select(-...1)

new <- read_csv("data/processed_data/new_processed.csv")

site_data <- read_csv("data/processed_data/site_data_processed.csv")
```

define a consistent palet, labels, and theme to be used in multiple figures
## still would be nice to have palet for harvested comparison
```{r}
# palet for comparisons of all 4 site types
pal <- brewer.pal(4, "BrBG")

# site labels for 4 site comparison, use in several plots
site_labels <-  c("Shade \n (n = 12)", "Sun \n (n = 11)", 
                              "Abandoned \n (n = 4)", "Forest \n (n = 5)")

# site labels for harvest, non-harvest comparison
labels_harvest <-  c("Not harvested \n (n = 9)", "Harvested \n (n = 23)")

# set a theme for ggplots
my_theme <- function() {
  theme_minimal() + 
   theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(size = 14, color = "gray25"),
        legend.text = element_text(size = 12))
}
```

## make a site_type dataframe - for downstream comparisons 
```{r}
site_type <- site_data %>% # create site_type df, just name of site and the type
  filter(Transecto == 1) %>% # so just one row per site
  select(Sitio, Tipo)

# reorder factor levels now for downstream plotting
site_type$Tipo <- factor(site_type$Tipo, c("N", "C", "V", "B"))

# here add harvested column for later analyses
site_type <- site_type %>%
  mutate(harvested = case_when(
    Tipo == "N" ~ "Yes",
    Tipo == "C" ~ "Yes",
    Tipo == "B" ~ "No",
    Tipo == "V" ~ "No")
    )
```






# RICHNESS ACROSS LANDUSE TYPES


richness by type - just number of species
```{r}
# group by site and type to retain type for ggplot
richness_by_site <- new %>%
  group_by(Sitio, Tipo) %>%
  summarize(species = list(sort(unique(`Final_ID`))),
            no_species = n_distinct(`Final_ID`))

# Group by land_use and calculate species richness
richness_summary <- richness_by_site %>%
  group_by(Tipo) %>%
  summarize(mean_species_richness = mean(no_species),
            median_species_richness = median(no_species),
            min_species_richness = min(no_species),
            max_species_richness = max(no_species))
```


## before selecting statistical test, need to look at dist. of richness by site
- use Poisson dist. b/c using count data with no upper limit!
```{r}
(richness_by_site_hist <- ggplot(richness_by_site, aes(no_species)) +
  geom_histogram())
```


## run glm for species richness across landuse types, using Poisson dist.
- bosque is sig different from nacional, viejo almost is!
- NEED these final IDs to come back!
```{r}
richness_by_type_glm <- glm(no_species ~ Tipo, data = richness_by_site, 
             family = poisson(link = "log"))
summary(richness_by_type_glm)

# here use DHARMa package to see if we're using the correct model
# no significant problems detected!
simulationOutput_richness_by_type <- simulateResiduals(
  fittedModel = richness_by_type_glm)
plot(simulationOutput_richness_by_type)
```


plot species richness across landuse types
```{r}
# reorder factor levels
richness_by_site$Tipo <- factor(richness_by_site$Tipo,
                         c("N", "C", "V", "B"))

# plot species richness across landuse types using palet, my_theme
plot_richness_by_site <- ggplot(richness_by_site, 
                                     aes(x = Tipo, y = no_species, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  # order needs to match factor levels above
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Landuse types",
       y = "Species richness",
       title = "Species richness across landuse types")
plot_richness_by_site
```


## harvested vs. non-harvested for richness

add harvest column (active plantation or not) to richness_by_site df
```{r}
richness_by_site <- richness_by_site %>%
  mutate(harvested = case_when( # add harvested column
    Tipo == "N" ~ "Yes",
    Tipo == "C" ~ "Yes",
    Tipo == "B" ~ "No",
    Tipo == "V" ~ "No")
    )
```

run glm for harvest on species richness using Poisson dist
- sig difference between harvested and non-harvested areas
```{r}
richness_by_harvest_glm <- glm(no_species ~ harvested, data = richness_by_site, 
             family = poisson(link = "log"))
summary(richness_by_harvest_glm)

# within group deviation from uniformity n.s.
# Levene Test for homogeneity of variance n.s.
simulationOutput_richness_by_harvest <- simulateResiduals(
  fittedModel = richness_by_harvest_glm)
plot(simulationOutput_richness_by_harvest)
```


B+V have sig higher richness than N+C
## aesthetically would want to mix colors from previous plot
```{r}
plot_richness_by_harvest <- ggplot(richness_by_site, 
                                   aes(x = harvested, y = no_species, fill = harvested)) +
  scale_x_discrete(labels = c(labels_harvest)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Active cacao plantation?", y = "Species richness") +
  ggtitle("Species richness in harvested and non-harvested areas")
plot_richness_by_harvest
```


## rarified species richness

## NEED TO GO THROUGH rarefy_scrap to piece this piece together

technique to estimate expected species richness
- allows calculation of species richness for given number of samples based on
  construction of rarefaction curves
  
issue is that the larger number of ind sampled, more species will be found
- rarefaction curves created by randomly re-sampling pool of N samples multiple
  times and then plotting average number of species found in each sample
- rarefaction generates expected number of species in small collection of n ind
  drawn from large pool of N samples
- typically grow rapidly at first, then slowly as only rarest species remain to
  be sampled

## First need to convert data into 3 column format
habitat, species, abundance for labdsv - use matrify!

new_mat - three column format to use in labdsv package
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)
```

use rarefy and rarecurve functions
```{r}
# ind for each site
site_abundance <- rowSums(new_mat)
site_abundance

# ind for each species
species_abundance <- colSums(new_mat)
species_abundance
```

## remove VC-C - use new_mat_abundant for rarefaction

rarefaction uses smallest number of obs per sample to extrapolate expected
number of all other samples only had that number of obs
- VC-C only had 6!!! may want to proceed without this  in the future
- next lowest is 22
```{r}
# including VC-C raremin is 6
rare_min <- min(rowSums(new_mat))
rare_min

rare_species_richness <- rarefy(new_mat, rare_min)
rare_species_richness

# it won't allow us to override this to 22 (makes sense)
# so will need to remove VC-C via row position
new_mat_abundant <- new_mat %>%
  filter(!row_number() %in% c(31))

# let's see what happens with this as our raremin, and only using 31/32 sites
rare_min_abundant <- min(rowSums(new_mat_abundant))
rare_min_abundant

# rarefy gives expected 'rarefied' number of species per site(not obs) if only 22 collected
rare_species_abundant <- rarefy(new_mat_abundant, rare_min_abundant)
rare_species_abundant
```

now join rare_species_abundant with site_type
## need to make sure that they're the same length!

## COME BACK HERE MAKE SURE MATCHES!!!!##
```{r}
# enframe rare_species_abundant which is currently a vector
rare_species_abundant_df <- enframe(rare_species_abundant)

# need to change name to Sitio to match for left_join
# also change value to Shannon_Index
colnames(rare_species_abundant_df) <- c("Sitio","Shannon_Index")

# need to make sure that site_type matches length
### HERE!!!!!!!! ####

# join by "Sitio" to site_type
#rare_species_abundant_df <- left_join(site_type, rare_species_abundant, by = "Sitio")
```












# ABUNDANCE ACROSS LANDUSE TYPES








number of new individuals 
## could eventually complicate this to look at pop size estimates?

abundance by site df
```{r}
# note this is same as site_abundance above
abundance_by_site <- new %>%
  group_by(Sitio, Tipo) %>%
  count()

abundance_by_site$Tipo <- factor(abundance_by_site$Tipo, c("N", "C", "V", "B"))
```


## before selecting statistical test, need to look at dist. of abundance by site
- use Poisson dist. b/c using count data with no upper limit!
```{r}
(abundance_by_site_hist <- ggplot(abundance_by_site, aes(n)) +
  geom_histogram(binwidth = 5))
```


## run glm for abundance across landuse types, using Gaussian dist.

DONT RUN POISSON HERE!
- quantile deviations detected
- combined adjusted quantile test significant

There is no sig difference in abundance across landuse types
```{r}
# default family dist is Gaussian
abundance_by_type_glm <- glm(n ~ Tipo, data = abundance_by_site)
summary(abundance_by_type_glm)

# no significant problems detected w/Gaussian!
simulationOutput_abundance_by_type <- simulateResiduals(
  fittedModel = abundance_by_type_glm)
plot(simulationOutput_abundance_by_type)
```

```{r}
plot_abundance_by_site <- ggplot(abundance_by_site, aes(x = Tipo, y = n, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Landuse type",
       y = "Individuals per site",
       title = "Abundance across landuse types")
plot_abundance_by_site
```












# DIVERSITY ACROSS LANDUSE TYPES













Look at how different diversity metrics compare across landuse types

## First need to convert data into 3 column format
habitat, species, abundance for labdsv - use matrify!

new_mat - three column format to use in labdsv package
```{r}
three_column <- new %>%
  group_by(Sitio) %>%
  count(`Final_ID`)

# needs to be converted to data.frame!
three_column <- data.frame(three_column)

new_mat <- matrify(three_column)
```


## Diversity Indices


use 3 column format to get diversity indices for each site!
```{r}
shannon <- diversity(new_mat, index = "shannon")
simpson <- diversity(new_mat, index = "simpson")
inv_simpson <- diversity(new_mat, index = "invsimpson")
```

join enframed shannon_df to site_type
```{r}
# enframe shannon which is currently a vector
shannon_df <- enframe(shannon)

# need to change name to Sitio in shannon_df to match for left_join
# also change value to Shannon_Index
colnames(shannon_df) <- c("Sitio","Shannon_Index")

# join by "Sitio"
shannon_df <- left_join(site_type, shannon_df, by = "Sitio")
```

do the same thing for simpson_df
```{r}
simpson_df <- enframe(simpson)

colnames(simpson_df) <- c("Sitio","Simpson_Index")

simpson_df <- left_join(site_type, simpson_df, by = "Sitio")
```

and inverse simpson_df
```{r}
inv_simpson_df <- enframe(inv_simpson)

colnames(inv_simpson_df) <- c("Sitio","Inv_Simpson_Index")

inv_simpson_df <- left_join(site_type, inv_simpson_df, by = "Sitio")
```


## look at dist of diversity metrics
```{r}
# shannon - clearly not normal dist! 
(shannon_hist <- ggplot(shannon_df, aes(Shannon_Index)) +
  geom_histogram(binwidth = .05))

# simpson - clearly not normal dist! 
(simpson_hist <- ggplot(simpson_df, aes(Simpson_Index)) +
  geom_histogram(binwidth = .05))

# inv_simpson - clearly not normal dist! 
(inv_simpson_hist <- ggplot(inv_simpson_df, aes(Inv_Simpson_Index)) +
  geom_histogram(binwidth = .1))
```

## Thought Poisson dist would work for these b/c not normal and from count data
- woah! sig deviation for Poisson dist.
- quantile deviations detected
- combined adjusted quantile test sig
## But instead use Gaussian dist! 

## sig higher shannon div in Bosque vs. other sites
```{r}
# shannon w/ Gaussian dist - all good
shannon_by_type_glm <- glm(Shannon_Index ~ Tipo, data = shannon_df)
summary(shannon_by_type_glm)

# no significant problems detected using Gaussian dist!
simulationOutput_shannon_by_type <- simulateResiduals(
  fittedModel = shannon_by_type_glm)
plot(simulationOutput_shannon_by_type)
```

## sig higher simpson div in Bosque vs. other sites
```{r}
# simpson w/ Gaussian dist
simpson_by_type_glm <- glm(Simpson_Index ~ Tipo, data = simpson_df)
summary(simpson_by_type_glm)

# no significant problems detected using Gaussian dist!
simulationOutput_simpson_by_type <- simulateResiduals(
  fittedModel = simpson_by_type_glm)
plot(simulationOutput_simpson_by_type)
```

## no sig difference in inv_simpson diversity across landtypes
# NOTE THAT HERE USING Inverse Gaussian Dist
```{r}
# inv_simpson w/ inverse gaussian dist.
inv_simpson_by_type_glm <- glm(Inv_Simpson_Index ~ Tipo, data = inv_simpson_df,
                               family = inverse.gaussian(link = "1/mu^2"))
summary(inv_simpson_by_type_glm)

# no significant problems detected using Gamma dist!
# all other tested dist threw errors w/ DHARMa
simulationOutput_inv_simpson_by_type <- simulateResiduals(
  fittedModel = inv_simpson_by_type_glm)
plot(simulationOutput_inv_simpson_by_type)

# Gaussian - quantile deviations detected - comb adjust - quant test sig
# Binomial - error - y values must be b/ween 0:1
# Gamma - deviation significant w/ outlier test but better than others
# Inv Gaussian - works! need to load additional cran package SuppDists
```

## plot how diversity indices vary across land types
- these are pretty much identical regardless of index used

shannon
```{r}
plot_shannon <- ggplot(shannon_df, 
                       aes(x = Tipo, y = Shannon_Index, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Landtype",
       y = "Shannon diversity",
       title = "Shannon diversity across landuse types")
plot_shannon
```

simpson
```{r}
plot_simpson <- ggplot(simpson_df, 
                       aes(x = Tipo, y = Simpson_Index, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() + 
  labs(x = "Landtype",
       y = "Simpson diversity",
       title = "Simpson diversity across landuse types")
plot_simpson
```

inv simpson
```{r}
plot_inv_simpson <- ggplot(inv_simpson_df, 
                           aes(x = Tipo, y = Inv_Simpson_Index, fill = Tipo)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c(site_labels)) +
  my_theme() +
  labs(x = "Ecological landtype",
       y = "Inverse simpson diversity",
       title = "Inverse Simpson diversity across landuse types")
plot_inv_simpson
```


# Compare diversity indices for harvested vs. unharvested sites

do the same thing as above, using harvested, unharvested column
- glms across harvested, unharvested

## sig higher shannon div in non-harvested sites
```{r}
# using Gaussian dist.
shannon_harvest_glm <- glm(Shannon_Index ~ harvested, data = shannon_df)
summary(shannon_harvest_glm)

# no issues with Gaussian dist
simulationOutput_shannon_harvest <- simulateResiduals(
  fittedModel = shannon_harvest_glm)
plot(simulationOutput_shannon_harvest)
```

## sig higher simpson div in non-harvested sites
```{r}
# using Gaussian dist.
simpson_harvest_glm <- glm(Simpson_Index ~ harvested, data = simpson_df)
summary(simpson_harvest_glm)

# no issues with Gaussian dist
simulationOutput_simpson_harvest <- simulateResiduals(
  fittedModel = simpson_harvest_glm)
plot(simulationOutput_simpson_harvest)
```

## no sig difference in inv_simpson div in harvested vs. non-harvested
# HERE THE GAUSSIAN DIST WORKS FINE FOR INV_SIMPSON - WHY?!?!?!
```{r}
# using Gaussian dist.
inv_simpson_harvest_glm <- glm(Inv_Simpson_Index ~ harvested, data = inv_simpson_df)
summary(inv_simpson_harvest_glm)

# no issues with Gaussian dist
simulationOutput_inv_simpson_harvest <- simulateResiduals(
  fittedModel = inv_simpson_harvest_glm)
plot(simulationOutput_inv_simpson_harvest)
```

now make figures for each
## aesthetically think about fill colors

shannon_harvest
```{r}
plot_shannon_harvest <- ggplot(shannon_df, 
                               aes(x = harvested, y = Shannon_Index, 
                                   fill = harvested)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Harvested",
       y = "Shannon diversity",
       title = "Shannon diversity across harvested and unharvested sites")
plot_shannon_harvest
```

simpson_harvest
```{r}
plot_simpson_harvest <- ggplot(simpson_df, 
                               aes(x = harvested, y = Simpson_Index, 
                                   fill = harvested)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Harvested",
       y = "Simpson diversity",
       title = "Simpson diversity across harvested and unharvested sites")
plot_simpson_harvest
```

inv_simpson_harvest
```{r}
plot_inv_simpson_harvest <- ggplot(inv_simpson_df, 
                               aes(x = harvested, y = Inv_Simpson_Index, 
                                   fill = harvested)) +
  geom_boxplot() +
  my_theme() +
  labs(x = "Harvested",
       y = "Inverse Simpson diversity",
       title = "Inverse Simpson diversity across harvested and unharvested sites")
plot_inv_simpson_harvest
```


















# COMPOSITION ACROSS LANDUSE TYPES....















## Pair-wise dissimilarity

Calculate pair-wise dissimilarity (distance) using vegdist
- gower and bray-curtis are good in detecting underlying ecological gradients
```{r}
par(mfrow = c(1, 2))
bray <- vegdist(new_mat, "bray") 
gower <- vegdist(new_mat, "gower")

hist(bray, xlim = range(0.0,1.0))
hist(gower, xlim = range(0.0,1.0))
```

## CONTINUE HERE
dissimilarity analysis is good way to explore variability in community comp
- next steps would be to do some sort of cluster analysis
- see where community associations exist
- but switching gears to look at rarefaction


## Non-metric Multidimensional Scaling


## need to reorder new_mat rows to allow nmds to draw borders of communities
- goal is to represent info from multiple dimensions into a few, so can visualize
  and interpret.
- so here, present the similarity of communities in 2-dimensional space

this is going to be ugly, but have triple checked against original new_mat
```{r}
new_order <- c(1,3,5,8,19,20,22,25,27,29,30,32, # shade (n=12)
               2,4,7,16,18,21,23,24,26,28,31, # sun (n=11)
               9,13,14,17, # abandoned (n=4)
               6,10,11,12,15 # forest (n=5)
               )
# Defined the desired order of rows: Shade, Sun, Abandoned, Forest

# Reorder rows
new_mat <- new_mat[new_order, ]
```



## create NMDS plot and draw boundaries of landuse types

code from Peat Clark's vegan tutorial
```{r}
new_mat_NMDS=metaMDS(new_mat, k=2)
# use community by species matrix from above

plot(new_mat_NMDS)
# here sites are open circles, species are red crosses

# here help visualize by labeling specific sites and species

# ordination plot function especially for congested plots
ordiplot(new_mat_NMDS,type="n")
# this function adds text or points to ordination plots
orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

orditorp(new_mat_NMDS,display="sites",cex=1.25,air=0.01)
```

## use ordihull to help visualize different treatments unifying sites

assign treatment levels per landuse type on reordered new_mat_reordered
```{r}
landuse_treatments=c(rep("Shade",12),
                     rep("Sun",11), 
                     rep("Abandoned",4), 
                     rep("Forest",5))
# ensure that labels are in correct order and have correct length

ordiplot(new_mat_NMDS,type="n")

ordihull(new_mat_NMDS,groups=landuse_treatments,draw="polygon",
         col="grey90",label=F)

orditorp(new_mat_NMDS,display="species",col="red",air=0.01)

# here colors need to align with landuse_treatments from above!
# of course need to use chocolate!
orditorp(new_mat_NMDS,display="sites",col=c(rep("chocolate4",12),
                                            rep("chocolate1",11),
                                            rep("lightgreen",4),
                                            rep("darkgreen",5)), 
         air=0.01,cex=1.25)
```

## very cool! see a lot of overlap in species simmilarity across landuse types
- but, forest and abandoned sites are similar, minus 3 outliers (FLL-B, FLL-C, FCAT-B3)
- these are also the lower elevation sites of the 9!
- think that elevation will be an important thing to include when comparing communities!
## CONTINUE HERE looking at elevation and group identity for NMDS

## NEXT STEPS: MAKE ORDINATION OFF GROUP, COMMUNITY DISSIMILARITY





