---
title: "rarefy_scrap"
author: "Holden Jones"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# RAREFACTION - FOLLOWED FROM RICHNESS ABOVE
# REVIEW ALL OF THIS - COPIED FROM RANAS_EXPLORE

technique to estimate expected species richness
- allows calculation of species richness for given number of samples based on
  construction of rarefaction curves
  
issue is that the larger number of ind sampled, more species will be found
- rarefaction curves created by randomly re-sampling pool of N samples multiple
  times and then plotting average number of species found in each sample
- rarefaction generates expected number of species in small collection of n ind
  drawn from large pool of N samples
- typically grow rapidly at first, then slowly as only rarest species remain to
  be sampled

use rarefy and rarecurve functions
```{r}
spAbund <- rowSums(new_mat)
spAbund

test <- colSums(new_mat)
test
```

rarefaction uses smallest number of obs per sample to extrapolate expected
number if all other samples only had that number of obs
- VC-C only had 3!!! may want to proceed without this  in the future
- GP-N also only 9, without these two the next lowest is 22
```{r}
# including VC-C raremin is 3
raremin <- min(rowSums(new_mat))
raremin

sRare <- rarefy(new_mat, raremin)
sRare

# it won't allow us to override this to 22 (makes sense)
# so will need to remove VC-C, GP-N from sites, via row position
new_mat_abundant <- new_mat %>%
  filter(!row_number() %in% c(19, 31))

# let's see what happens with this as our raremin, and only using 30/32 sites
raremin_abundant <- min(rowSums(new_mat_abundant))
raremin_abundant

# sRare gives expected 'rarefied' number of species (not obs) if only 22 collected
sRare_abundant <- rarefy(new_mat_abundant, raremin_abundant)
sRare_abundant

rare_shannon_aov <- aov(value ~ Tipo, data = sRare_abundant_df)
summary(rare_shannon_aov)
```

## no statistical difference in rarefied species richness across types
*note that using rarefaction to remove effects of diff sample sizes is bad!*
```{r}
# join tipo data with sRare_abundant in a df
sRare_abundant_df <- sRare_abundant %>% 
  enframe() %>% 
  full_join(site_type, by = c("name" = "Sitio"))

sRare_abundant_aov <- aov(value ~ Tipo, data = sRare_abundant_df)
summary(sRare_abundant_aov)
# p-value is 0.183 (04/9/23) - check back in when sequencing is done!
```

visualization using rarecurve, for ggplot?
## make rarecurve for each type of site
## stuck!
```{r}
# rarecurve for each site, would be nice to have for each type!
rarecurve(new_mat_abundant, col = "blue", cex = 0.4)
```